---
title: "DataCORE"
subtitle: "Filtering, Subsetting & Group Summaries in Pandas"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE)
```

```{python py-setup, include=FALSE}
import pandas as pd
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Filtering & Subsetting &mdash; Asking Your Data a Question</h4></li>
    <li><h4>Handling Missing Values &amp; Group-Level Summaries</h4></li>
    <li><h4>Scientific Storytelling &mdash; The One-Number Story</h4></li>
</ol>

.font90[*This morning you named a target population and a DV. This afternoon we make the data answer for it.*]

---
class: segue-green

# 1:40 &ndash; 2:45 &middot; Filtering & Subsetting
## .white[Core Techniques Every Researcher Needs]

---
### `r fontawesome::fa("filter", height = "1em", fill = "#2e8b57")` Why filter? From the *whole* dataset to *your* question

.pull-left[
.font90[
- A dataset holds **everyone**. Your research question is about a **specific group**.

- This morning: maybe your population is **9th graders**, or **students who sleep less than 7 hours**.

- **Filtering** keeps only the rows that match &mdash; so every later number describes the group you actually care about.
]
]

.pull-right[
.content-box-blue[
.font90[**Big idea:** a filter is a **yes/no test** applied to every row. Rows that pass stay; rows that fail drop out.]
]
]

.font80[`r fontawesome::fa("table", fill = "#4a7fb5")` &nbsp; We continue with `study_habits.csv` &mdash; swap in **your** team file.]

---
### `r fontawesome::fa("file-csv", height = "1em", fill = "#2e8b57")` Reload the dataset

.panelset[
.panel[.panel-name[Python]
```{python}
import pandas as pd
df = pd.read_csv("study_habits.csv")   # rows = students, columns = variables   #<<
print(df.head())
```
]
.panel[.panel-name[R]
```{r}
library(tidyverse)
df <- read_csv("study_habits.csv")     # rows = students, columns = variables   #<<
head(df)
```
]
]

.font80[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; Columns: `grade_level`, `study_method`, `hours_of_sleep`, `test_score`.]

---
### `r fontawesome::fa("code-branch", height = "1em", fill = "#2e8b57")` The mask: a yes/no test for every row

.font90[Write the **condition** alone first &mdash; it returns `True`/`False` for each row.]

.panelset[
.panel[.panel-name[Python]
```{python}
mask = df["hours_of_sleep"] < 7   # a Boolean Series: one True/False per row   #<<
print(mask.head())
```
]
.panel[.panel-name[R]
```{r}
mask <- df$hours_of_sleep < 7     # a logical vector: one TRUE/FALSE per row    #<<
head(mask)
```
]
]

.font80[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; The mask is the **question**; the filter is the **answer**.]

---
### `r fontawesome::fa("filter", height = "1em", fill = "#2e8b57")` `df[df['col'] < value]` &mdash; filter to a subpopulation

.panelset[
.panel[.panel-name[Python]
```{python}
under_7 = df[df["hours_of_sleep"] < 7]   # keep only rows that pass the test    #<<
print(under_7)
```
]
.panel[.panel-name[R]
```{r}
under_7 <- df %>% filter(hours_of_sleep < 7)   # keep only rows that pass        #<<
under_7
```
]
]

.font80[`r fontawesome::fa("equals", fill = "#4a7fb5")` &nbsp; Use `==` for categories: `df[df["grade_level"] == "9th"]`.]

---
### `r fontawesome::fa("layer-group", height = "1em", fill = "#2e8b57")` Combine conditions with `&` (and) / `|` (or)

.panelset[
.panel[.panel-name[Python]
```{python}
# 9th graders who slept less than 7 hours -- wrap each condition in ()   #<<
subset = df[(df["grade_level"] == "9th") & (df["hours_of_sleep"] < 7)]
print(subset)
```
]
.panel[.panel-name[R]
```{r}
subset <- df %>% filter(grade_level == "9th", hours_of_sleep < 7)   #<<
subset
```
]
]

.content-box-yellow[
.font80[**Watch out:** in Python use `&`/`|` (not `and`/`or`) and **parentheses** around each condition.]
]

---
### `r fontawesome::fa("triangle-exclamation", height = "1em", fill = "#2e8b57")` `df.dropna()` &mdash; handle missing values *intentionally*

.font90[Our data has blanks in `hours_of_sleep` and `test_score`. Drop rows missing a **key column** &mdash; not the whole table.]

.panelset[
.panel[.panel-name[Python]
```{python}
print("Before:", len(df))
clean = df.dropna(subset=["test_score"])   # drop rows with no DV value          #<<
print("After: ", len(clean))
```
]
.panel[.panel-name[R]
```{r}
cat("Before:", nrow(df), "\n")
clean <- df %>% drop_na(test_score)        # drop rows with no DV value          #<<
cat("After: ", nrow(clean), "\n")
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; `subset=` matters: dropping *any* missing value can throw away rows you needed.]

---
### `r fontawesome::fa("calculator", height = "1em", fill = "#2e8b57")` Compute the mean of your DV *for the subpopulation*

.panelset[
.panel[.panel-name[Python]
```{python}
ninth = df[df["grade_level"] == "9th"]      # filter to the population           #<<
print(ninth["test_score"].mean())           # mean of the DV                      #<<
```
]
.panel[.panel-name[R]
```{r}
ninth <- df %>% filter(grade_level == "9th")   # filter to the population         #<<
mean(ninth$test_score, na.rm = TRUE)            # mean of the DV                   #<<
```
]
]

.font80[`r fontawesome::fa("brain", fill = "#4a7fb5")` &nbsp; `na.rm` / pandas skips missing values automatically &mdash; know what you skipped.]

---
### `r fontawesome::fa("object-group", height = "1em", fill = "#2e8b57")` `df.groupby('category').mean()` &mdash; all groups at once

.panelset[
.panel[.panel-name[Python]
```{python}
df.groupby("study_method")["test_score"].mean()   # one mean per category         #<<
```
]
.panel[.panel-name[R]
```{r}
df %>% group_by(study_method) %>%                  # one mean per category         #<<
  summarise(mean_score = mean(test_score, na.rm = TRUE))
```
]
]

.font80[`r fontawesome::fa("chart-column", fill = "#4a7fb5")` &nbsp; `groupby` does the filtering for *every* group in one line &mdash; instant comparison.]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` Team Challenge
## .white[Filter to Your Population &middot; Compute Your DV]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Team Challenge &mdash; ~25 min

.pull-left[
.font90[
**As a team, write code that:**

<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">1.</strong> Filters **your** dataset to your **target population**.
</div>

<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">2.</strong> Drops rows missing your **DV** (`dropna`).
</div>

<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">3.</strong> Computes the **mean of your DV**.
</div>

**Instructor checks your output.**
]
]

.pull-right[
.content-box-blue[
.font90[**Then discuss as a team:**

What does this number actually mean **in real terms**?

*Not* "the mean is 78" &mdash; but *"the average test score among 9th graders who slept under 7 hours was 78 out of 100."*]
]
]

.font80[`r fontawesome::fa("people-group", fill = "#4a7fb5")` &nbsp; Record your number **and its plain-English meaning** on your Mission Sheet.]

---
### `r fontawesome::fa("robot", height = "1em", fill = "#2e8b57")` AI prompt practice &mdash; then break it

.pull-left[
.font90[
**1.** Ask your AI tool:

.content-box-blue[.font80[*"Write Python code to filter my DataFrame to rows where **[column]** equals **[value]**."*]]

**2.** **Run** the code on your real data.

**3.** **Find the edge case it misses.**
]
]

.pull-right[
.content-box-yellow[
.font85[
**Edge cases AI often forgets:**

- Text vs. number: `"9th"` vs `9`
- Capitalization: `"Group"` vs `"group"`
- Hidden spaces: `"9th "`
- Missing values silently dropped
- `=` (assignment) vs `==` (test)
]
]
]

.font80[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; AI writes code fast &mdash; **you** are the one who checks it against the data.]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[The One-Number Story]

---
### `r fontawesome::fa("comment-dots", height = "1em", fill = "#2e8b57")` Report your DV mean in *one* sentence

.pull-left[
.font90[
Each team reports their filtered DV mean to the class in **one sentence**:
]

.content-box-green[
.font100[*"Among **[population]**, the average **[DV]** was **[X]**."*]
]

.content-box-blue[
.font80[*"Among **9th graders**, the average **test score** was **79.6**."*]
]
]

.pull-right[
.font90[
**As a class, discuss:**

- Does this number **surprise** you?

- Does it **match** what the literature found this morning?
]

.content-box-yellow[
.font85[`r fontawesome::fa("seedling", fill = "#4a7fb5")` &nbsp; This one sentence is the **seed of your Results section.**]
]
]

.font80[`r fontawesome::fa("people-arrows", fill = "#4a7fb5")` &nbsp; A good result is **one clear number** the audience can repeat back to you.]

---
class: segue-green

# .white[From a number to a finding.]
## .white[Tomorrow: comparing groups & visualizing.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-03-from-curiosity-to-research-question/pm/03-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-03-from-curiosity-to-research-question/pm/03-pm-datacore.pdf")
to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-03-from-curiosity-to-research-question/pm/03-pm-datacore.html")
```
