---
title: "Research Methods"
subtitle: "Day 12 · Session 12: Research Ethics, Final Slide Polish & References"
# author: "Hussain Hadah (he/him)"
date: "July 14, 2026"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | July 14, 2026</div>

<!--- `r rmarkdown::metadata$subtitle` | July 14, 2026-->

---

class: title-slide
background-position: 10% 90%, 100% 50%
background-size: 160px, 50% 100%
background-color: #0148A4

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

    .img-row {
      display: flex;
      gap: 1em;
      align-items: center;
      justify-content: center;
      margin-top: 0.8em;
    }
    .img-row img {
      max-width: 100%;
      border-radius: 6px;
    }
    .img-row .cell {
      flex: 1;
      text-align: center;
    }

    .remark-code {
      font-size: 0.75em !important;
    }

    .card-grid {
      display: flex;
      flex-wrap: wrap;
      gap: 0.8em;
      margin-top: 0.6em;
    }

```

# .text-shadow[.white[Outline for Today]]

<ol>
    <li><h4 class="white">Research Ethics &mdash; Why It Matters</h4></li>
    <li><h4 class="white">'Cite That Right' Team Activity</h4></li>
    <li><h4 class="white">Slide Design Workshop</h4></li>
    <li><h4 class="white">Full Run-Through</h4></li>
</ol>

---
class: segue-yellow

# Session 12: Research Ethics, <br> Slide Polish & References `r top_icon("gavel")`

---
### Learning Objectives

.font90[
#### `r fontawesome::fa("scale-balanced", height = "1em", fill = "#4a7fb5")` &nbsp; Understand core principles of research ethics
Attribution, honesty, limitations, and consent &mdash; the values that make research trustworthy.

#### `r fontawesome::fa("quote-right", height = "1em", fill = "#4a7fb5")` &nbsp; Complete the References slide with proper citations
Every source cited, every citation consistent, AI use disclosed.

#### `r fontawesome::fa("wand-magic-sparkles", height = "1em", fill = "#4a7fb5")` &nbsp; Polish slides for visual clarity and professional appearance
One idea per slide, clean design, findings-first chart titles.

#### `r fontawesome::fa("person-chalkboard", height = "1em", fill = "#4a7fb5")` &nbsp; Practice a full presentation run-through for the first time
End to end, timed, with focused peer and instructor feedback.
]

---
class: segue-yellow

# Morning: ResearchCORE `r top_icon("sun")`

---
class: segue-green

# Research Ethics &mdash; <br> Why It Matters `r top_icon("balance-scale")`

---
### Why Ethics Matters

.pull-left[.font90[
.content-box-yellow[
**Research earns trust only when it is done honestly.** One dishonest study can undo the credibility of an entire field.
]

As a researcher you make promises &mdash; to your participants, your readers, and the people your findings might affect.

**Four pillars we'll return to today:**

- **Attribution** &mdash; credit every source and tool
- **Honesty** &mdash; report what you found, not what you hoped
- **Limitations** &mdash; own what your study can't say
- **Consent** &mdash; respect the people behind the data
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` The stakes are real**

Findings become policy, medicine, and code. Someone will act on what you publish &mdash; ethics is how you make sure that action is well-founded.
]
]]

---
### Core Ethical Considerations

<div class="card-grid">
  <div class="content-box-blue" style="flex:1 1 45%;">
    <p><strong>`r fontawesome::fa("user-shield", fill = "#4a7fb5")` Respect Privacy</strong></p>
    <p class="font85">Protect participants' identities and personal information. Use anonymous responses whenever possible.</p>
  </div>
  <div class="content-box-blue" style="flex:1 1 45%;">
    <p><strong>`r fontawesome::fa("file-signature", fill = "#4a7fb5")` Informed Consent</strong></p>
    <p class="font85">Ensure participants understand what they are agreeing to. Get permission from parents for minors.</p>
  </div>
  <div class="content-box-green" style="flex:1 1 45%;">
    <p><strong>`r fontawesome::fa("hand-holding-heart", fill = "#4a7fb5")` Minimize Harm</strong></p>
    <p class="font85">Design research that does not cause physical or emotional distress to participants.</p>
  </div>
  <div class="content-box-green" style="flex:1 1 45%;">
    <p><strong>`r fontawesome::fa("stamp", fill = "#4a7fb5")` Get Approvals</strong></p>
    <p class="font85">Obtain necessary permissions from teachers, administrators, or other authorities.</p>
  </div>
</div>

---
### Plagiarism & Proper Attribution

.pull-left[.font85[
.content-box-yellow[
**Plagiarism** = using someone's words, ideas, data, or figures without credit &mdash; whether on purpose or by accident.
]

**Give credit for:**
- Direct quotes *and* paraphrased ideas
- Data, tables, and figures you didn't create
- Methods or frameworks you borrowed
]]

.pull-right[.font85[
.content-box-green[
**`r fontawesome::fa("check", fill = "#009E73")` The safe habit**

When in doubt, cite. A citation never weakens your work &mdash; it shows you know the landscape and stand on solid ground.
]

.content-box-blue[
**`r fontawesome::fa("quote-left", fill = "#4a7fb5")` Quote vs. paraphrase**

Quoting copies exact words (in quotation marks); paraphrasing restates in your own words. **Both** still need a citation.
]
]]

---
### AI Ethics &mdash; What Do You Have to Disclose?

.pull-left[.font85[
Using AI tools is allowed &mdash; **hiding it is not.** Major journals (Nature, Science, APA) now require disclosure.

**Disclose:**
- Which AI tools you used
- What you used them *for* (coding, editing, brainstorming)
- That you **verified** the outputs yourself
]]

.pull-right[.font85[
.content-box-blue[
**`r fontawesome::fa("robot", fill = "#4a7fb5")` A sample attribution statement**

*"AI tools were used in this project for code drafting and language editing. All analyses were conducted and verified by the research team. All interpretations and conclusions are our own."*
]

.content-box-yellow[
.font90[`r fontawesome::fa("circle-exclamation", fill = "#4a7fb5")` &nbsp; **You** remain responsible for every number and claim &mdash; the AI is a tool, not an author.]
]
]]

---
### Researcher Bias &mdash; and How to Minimize It

.pull-left[.font90[
**Bias** is the pull toward the answer you *want* to find. Everyone has it; good researchers guard against it.

**Common forms:**
- **Confirmation bias** &mdash; noticing only evidence that fits
- **Selection bias** &mdash; a sample that isn't representative
- **Funding/interest bias** &mdash; a stake in the outcome
]]

.pull-right[.font90[
.content-box-green[
**`r fontawesome::fa("shield-halved", fill = "#4a7fb5")` Guardrails**

- Pre-decide your analysis *before* seeing results
- Report findings that **don't** fit your hypothesis
- Show your full data and code
- Invite someone to challenge your interpretation
]
]]

---
### Overstating Findings: p-Hacking & Cherry-Picking

.font80[
.pull-left[
.content-box-yellow[
**`r fontawesome::fa("magnifying-glass-dollar", fill = "#c08a00")` p-hacking**

Trying test after test until something crosses *p* &lt; 0.05, then reporting only that one. It manufactures "significance" that won't replicate.
]
]

.pull-right[
.content-box-yellow[
**`r fontawesome::fa("hand-pointer", fill = "#c08a00")` Cherry-picking**

Reporting only the results, cases, or time-windows that support your story &mdash; and quietly dropping the rest.
]
]
]

.content-box-green[
.font85[`r fontawesome::fa("check", fill = "#009E73")` &nbsp; **The honest alternative:** report all the analyses you ran, use hedged language (*suggests*, *is associated with*), and let a non-significant result stand as a finding. A true null is not a failure.]
]

---
### The Responsibility of Publishing

.pull-left[.font90[
Once your findings are public, **you no longer control how they're used.**

Ask before you publish:

- Who might **act** on this result?
- Could it be **misread** or **misused** to harm a group?
- Have I stated my **limitations** loudly enough to prevent overreach?
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("bullhorn", fill = "#4a7fb5")` Frame findings carefully**

A single correlation, stripped of caveats, can become a headline or a policy. Precise, honest framing is part of your ethical duty &mdash; not an afterthought.
]
]]

---
### When Research Goes Wrong

.three-col[
.col[
`r fontawesome::fa("user-doctor", height = "2.2em", fill = "#D55E00")`

**Tuskegee Study**

For 40 years, treatment was withheld from Black men with syphilis without their informed consent &mdash; the case that reshaped U.S. consent rules.
]
.col[
`r fontawesome::fa("flask-vial", height = "2.2em", fill = "#D55E00")`

**Data Fabrication**

Fabricated results in psychology helped fuel the **replication crisis** &mdash; famous findings that collapsed when others tried to reproduce them.
]
.col[
`r fontawesome::fa("camera", height = "2.2em", fill = "#D55E00")`

**AI Facial Recognition Bias**

Systems trained on skewed data misidentified women and darker-skinned faces at far higher rates &mdash; biased data, biased harm.
]
]

--

.content-box-yellow[
.font85[**The common thread:** each failure came from ignoring one of our pillars &mdash; consent, honesty, or attribution of the data's limits.]
]

---
class: segue-green

# 'Cite That Right' <br> Team Activity `r top_icon("quote-right")`

---
### 'Cite That Right' &mdash; Team Activity (20 min)

.pull-left[.font90[
Complete the **'Cite That Right' Handout** for your full reference list. Run these four checks:

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text">Every source <strong>cited in your slides</strong> appears in References.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">Every source <strong>in References</strong> was actually cited in a slide.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text"><strong>Consistent citation style</strong> throughout (pick one and hold it).</div></div>
<div class="chev-row-sm"><div class="chev-sm">4</div><div class="chev-text"><strong>AI tools documented</strong> with an attribution statement.</div></div>
]]

.pull-right[.font80[
.content-box-blue[
**`r fontawesome::fa("list-check", fill = "#4a7fb5")` In-text &rarr; reference list**

*In a slide:* "…failing to plan is planning to fail (Yin & Ogut, 2025)."

*In References:*
Yin & Ogut (2025). *Nexellence ResearchCORE: An Introduction to Research Methods.* Nexellence Press.

Toska, G. (2023). *Counseling Centers in Türkiye.* National Conference on Counseling Psychology, Istanbul.
]
]]

.center[`r timer(20L)`]

---
class: segue-yellow

# Slide Design Workshop `r top_icon("palette")`

---
### Slide Design Principles

.pull-left[.font85[
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("lightbulb", fill = "#0148A4")`</div><div class="chev-text"><strong>One idea per slide.</strong> If a slide makes two points, split it.</div></div>
<div class="chev-row-sm"><div class="chev-sm">6&times;6</div><div class="chev-text"><strong>The 6&times;6 rule:</strong> at most 6 bullets, 6 words each. Slides are cues, not scripts.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("chart-column", fill = "#0148A4")`</div><div class="chev-text"><strong>Chart titles are findings,</strong> not labels: "Wages rise with experience" &gt; "Wage vs. experience".</div></div>
]]

.pull-right[.font85[
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("font", fill = "#0148A4")`</div><div class="chev-text"><strong>Two fonts, maximum.</strong> One for headings, one for body.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("circle-half-stroke", fill = "#0148A4")`</div><div class="chev-text"><strong>High-contrast colors.</strong> Dark text on light, or light on dark &mdash; never mid-gray on white.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("universal-access", fill = "#0148A4")`</div><div class="chev-text"><strong>Accessible palettes.</strong> Color-blind-safe; never rely on color alone to carry meaning.</div></div>
]]

--

.content-box-yellow[
.font85[`r fontawesome::fa("eye", fill = "#c08a00")` &nbsp; **Read every slide from the back of the room.** If you can't, your audience can't either.]
]

---
### The Slide Design Checklist

.pull-left[.font85[
Trade slides with another team and score each one:

.content-box-blue[
- One clear idea, stated in the title
- &le; 6 bullets, &le; 6 words each
- Chart title states the **finding**
- Axis labels legible at slide size
- &le; 2 fonts; consistent throughout
- High-contrast, color-blind-safe
- Sources cited on the slide
]
]]

.pull-right[.font85[
.content-box-green[
**`r fontawesome::fa("users", fill = "#009E73")` Peer review, out loud**

For each slide your partner team can't instantly "get," ask: *what is the one thing this slide is trying to say?* If it takes more than a sentence to answer, the slide needs work.
]

.content-box-yellow[
.font90[`r fontawesome::fa("scissors", fill = "#c08a00")` &nbsp; When in doubt, **cut**. Less on the slide means more attention on you.]
]
]]

---
### Slide Showdown

.font90[Put your work up for a friendly, fast critique:]

.pull-left[
.content-box-green[
**`r fontawesome::fa("thumbs-up", fill = "#009E73")` Show 2 slides of your *best* work**

What makes them land? Name the design choice so the class can steal it.
]
]

.pull-right[
.content-box-yellow[
**`r fontawesome::fa("circle-question", fill = "#c08a00")` Show 2 slides you're *unsure* about**

The class suggests one concrete fix each. No slide leaves without an action.
]
]

.center[`r timer(30L)`]

---
### Reference: The Presentation Outline

.font70[
| Slide | What it contains |
|---|---|
| **Title** | Project title, presenters, date, institution, logos |
| **Topic** | The global challenge; define it and say why it matters (cite sources) |
| **Literature Review** | What's known, arranged by theme (cite sources) |
| **Research Question** | Your question(s) and, if possible, a hypothesis |
| **Methods** | Secondary analysis; data source & year; sample size & traits; key variables; tools |
| **Results** | Descriptive stats per variable, then figures/tables (may be several slides) |
| **Discussion** | How you interpret what you found |
| **Conclusion** | Final takeaway and close |
| **References** | Full list, one consistent style |
]

.content-box-blue[
.font85[`r fontawesome::fa("route", fill = "#4a7fb5")` &nbsp; Every polished deck follows this arc. If a slide doesn't fit the arc, ask whether it belongs.]
]

---
class: segue-green

# Full Run-Through `r top_icon("chalkboard-teacher")`

---
### Full Run-Through (40 min)

.pull-left[.font90[
Each team presents the **entire** presentation, beginning to end &mdash; the first real rehearsal.

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("stopwatch", fill = "#0148A4")`</div><div class="chev-text"><strong>Time it:</strong> target <strong>10&ndash;12 minutes</strong> per team.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("volume-xmark", fill = "#0148A4")`</div><div class="chev-text"><strong>Class observes silently</strong> &mdash; take notes, hold questions.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("comment-dots", fill = "#0148A4")`</div><div class="chev-text"><strong>3 notes per team</strong> from the instructor.</div></div>
]]

.pull-right[.font90[
.content-box-green[
**`r fontawesome::fa("plus", fill = "#009E73")` 2 strengths**

What worked &mdash; keep doing it.
]
.content-box-yellow[
**`r fontawesome::fa("wrench", fill = "#c08a00")` 1 specific improvement**

One concrete, actionable change for next time &mdash; not a vague "be clearer."
]
]]

.center[`r timer(12L)`]

---
### What to Listen For

.three-col[
.col[
`r fontawesome::fa("clock", height = "2em", fill = "#4a7fb5")`

**Timing**

Does it fit 10&ndash;12 minutes? Where does it drag or rush?
]
.col[
`r fontawesome::fa("arrows-left-right-to-line", height = "2em", fill = "#4a7fb5")`

**Transitions**

Do the hand-offs between speakers and sections feel smooth?
]
.col[
`r fontawesome::fa("bullseye", height = "2em", fill = "#4a7fb5")`

**Takeaway**

Could you state the team's main finding in one sentence afterward?
]
]

--

.content-box-blue[
.font85[`r fontawesome::fa("seedling", fill = "#4a7fb5")` &nbsp; A first run-through is *supposed* to be rough. Its job is to surface the fixes while there's still time to make them.]
]

---
### Session 12: What You Can Do Now

.font80[
**Research ethics:**

- Explain the four pillars &mdash; **attribution, honesty, limitations, consent** &mdash; and spot p-hacking, cherry-picking, and researcher bias in your own work.
- Name why historical cases (Tuskegee, fabrication in the replication crisis, AI facial-recognition bias) still shape research rules today.

**References & attribution:**

- Complete a **References slide** where every citation matches a source in the text, in one consistent style, with AI use disclosed.

**Slide polish & delivery:**

- Apply the design principles &mdash; one idea per slide, the 6&times;6 rule, findings-first chart titles, &le; 2 fonts, high-contrast accessible color &mdash; and review a peer's deck with the checklist.
- Deliver a full, timed run-through and turn feedback (2 strengths + 1 fix) into concrete revisions.
]

```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)
# https://hhadah.github.io/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-12-research-ethics-slide-polish-references/am/12-class.html

to_pdf(from = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-12-research-ethics-slide-polish-references/am/12-class.html",
       to = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-12-research-ethics-slide-polish-references/am/12-class.pdf")
to_pdf(from = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-12-research-ethics-slide-polish-references/am/12-class.html")
```
