---
title: "Research Methods"
subtitle: "Day 5: Research Design & Methodology"
# author: "Hussain Hadah (he/him)"
date: "July 3, 2026"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | July 3, 2026</div>

<!--- `r rmarkdown::metadata$subtitle` | `r format(Sys.time(), '%d %B %Y')`-->

---

class: title-slide
background-position: 10% 90%, 100% 50%
background-size: 160px, 50% 100%
background-color: #0148A4

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
# Helper: a centered, inline countdown timer (click to start/pause/reset)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.6em auto;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2.2rem",
    warn_when = 30L,
    ...
  )
}
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

```

# .text-shadow[.white[Outline for Today]]

<ol>
    <li><h4 class="white">Team Presentations: Literature Findings</h4></li>
    <li><h4 class="white">Evidence of Importance</h4></li>
    <li><h4 class="white">Research Design 101: Choosing a Method</h4></li>
    <li><h4 class="white">Design Your Study</h4></li>
    <li><h4 class="white">Scientific Storytelling: Methods Paragraph</h4></li>
</ol>

---
class: segue-yellow

# Welcome to Day 5! `r top_icon("star")`

---
### Learning Objectives for Today

.font90[
.pull-left[
.content-box-blue[
**`r fontawesome::fa("compass-drafting", fill = "#4a7fb5")` Understand key research designs**

Experimental, quasi-experimental, observational, survey, and secondary data analysis.
]

.content-box-blue[
**`r fontawesome::fa("database", fill = "#4a7fb5")` Why secondary data?**

Clarify why the camp project uses **secondary data analysis** &mdash; and what its strengths and limitations are.
]
]

.pull-right[
.content-box-yellow[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` Spot limitations & bias**

Identify potential limitations and sources of bias in your planned study.
]

.content-box-green[
**`r fontawesome::fa("link", fill = "#4a7fb5")` Connect method to question**

Connect methodology choices to your ability to answer your research question.
]
]
]

---
class: segue

# Team Presentations: Literature Findings `r top_icon("users")`

---
### Team Presentations: Literature Findings

.pull-left[
```{r img_design, echo=FALSE, warning=FALSE, out.width="92%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-design.png")
grid.raster(img)
```
]

.pull-right[
<br>

#### `r fontawesome::fa("microphone", height = "1em", fill = "#4a7fb5")` &nbsp; Each team presents **1&ndash;2 interesting findings**
From your literature review (Session 4 Mission Sheet, Part 2).

<br>

#### `r fontawesome::fa("comment-dots", height = "1em", fill = "#4a7fb5")` &nbsp; Class gives **one** piece of feedback per team
> *"One question this raises for me..."*

<br>

.center[`r timer(30L)`]
]

---
### Team Activity: Evidence of Importance

.pull-left[
.content-box-blue[
**`r fontawesome::fa("clipboard-list", fill = "#4a7fb5")` Your task**

Using the **"Evidence of Importance"** Handout, compile **2&ndash;3 statistics or findings** from your literature review that show **why your research topic matters**.
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("bullhorn", fill = "#4a7fb5")` Why it matters**

This becomes the **opening of your final presentation** &mdash; the hook that convinces your audience the question is worth asking.
]
]

<div style="clear: both;"></div>

.center[.font90[*Work as a team*]]

.center[`r timer(20L)`]

---
class: segue-green

# Research Design 101 `r top_icon("drafting-compass")`

---
### Designing a Research Project

.pull-left[
```{r img_design2, echo=FALSE, warning=FALSE, out.width="95%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-design.png")
grid.raster(img)
```
]

.pull-right[
<br>

<br>

## `r fontawesome::fa("lightbulb", height = "1em", fill = "#4a7fb5")` Let's explore how to **plan, execute, and analyze** research that answers meaningful questions.

<br>

.font90[To make designing easier, follow our **Research Design Checklist** as a team.]
]

---
### Failing to Plan is Planning to Fail

.pull-left[
```{r img_plan, echo=FALSE, warning=FALSE, out.width="88%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-plan.png")
grid.raster(img)
```
]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("signs-post", fill = "#4a7fb5")` Research Requires Direction**

Like baking without a recipe or traveling without a map, research without planning leads to confusion.
]

.content-box-blue[
**`r fontawesome::fa("pen-ruler", fill = "#4a7fb5")` Create a Blueprint**

A well-designed project saves time and increases the chances of valid, meaningful findings.
]

.content-box-green[
**`r fontawesome::fa("chess", fill = "#4a7fb5")` Develop a Game Plan**

We'll cover methodologies, tools, variables, and ethical considerations for your research journey.
]
]]

---
### Choosing a Research Method

.font110[The first step in designing your project is selecting the **right method** to answer your question.]

<br>

.font90[**Methodology includes...**]

<div class="chev-flow">
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("route", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Your overall approach / strategy</strong><br>
    The big-picture plan for how you'll answer your question.</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("toolbox", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Your data collection tools / techniques</strong><br>
    The specific ways you'll gather your evidence.</p>
  </div>
</div>

---
### Approach: Quantitative vs. Qualitative

.pull-left[
.content-box-blue[
### `r fontawesome::fa("calculator", height = "1em", fill = "#4a7fb5")` Quantitative

**Focus on numbers and measurable data.**

Good for questions about *"how many,"* *"how much,"* or **testing relationships** between variables.
]
]

.pull-right[
.content-box-green[
### `r fontawesome::fa("comments", height = "1em", fill = "#4a7fb5")` Qualitative

**Focus on descriptions and non-numerical insights.**

Good for **exploring** experiences, opinions, and understanding *"why"* or *"how."*
]
]

<div style="clear: both;"></div>

.center[.font110[Your approach can be **Qualitative**, **Quantitative**, or **Mixed**.]]

---
### Where Does Your Data Come From?

.pull-left[
```{r img_methods, echo=FALSE, warning=FALSE, out.width="90%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-methods.png")
grid.raster(img)
```
]

.pull-right[
.content-box-blue[
**`r fontawesome::fa("box-archive", fill = "#4a7fb5")` Use already-collected data**

**Secondary Research** &mdash; analyze data others have gathered.

.font90[*This is what your camp project will use!*]
]

.content-box-yellow[
**`r fontawesome::fa("magnifying-glass-chart", fill = "#4a7fb5")` Collect your own data**

Use one or more methods &mdash; for example, run an **experiment** and/or make **natural observations**.
]
]

---
### Data Collection Tools & Techniques

.font90[
.pull-left[
.content-box-blue[
**`r fontawesome::fa("flask-vial", fill = "#4a7fb5")` Experiment**

Manipulate variables to test cause and effect.
]

.content-box-blue[
**`r fontawesome::fa("square-poll-vertical", fill = "#4a7fb5")` Surveys**

Collect responses from many people quickly.
]

.content-box-blue[
**`r fontawesome::fa("comment", fill = "#4a7fb5")` Interview**

One-on-one conversations for depth.
]
]

.pull-right[
.content-box-blue[
**`r fontawesome::fa("people-group", fill = "#4a7fb5")` Focus Groups**

Guided group discussions for multiple views.
]

.content-box-blue[
**`r fontawesome::fa("eye", fill = "#4a7fb5")` Observation**

Watch behavior as it naturally occurs.
]

.content-box-blue[
**`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` Case Study**

Deep dive into a single case or a few cases.
]
]
]

---
### Experiments: Testing Cause and Effect

.pull-left[.font80[
.content-box-blue[
**`r fontawesome::fa("circle-question", fill = "#4a7fb5")` When to Use**

- You can **manipulate** variables
- You need to **control** other factors
- Your question starts with *"What is the effect of...?"*
]
]]

.pull-right[.font80[
.content-box-green[
**`r fontawesome::fa("gears", fill = "#4a7fb5")` Key Components**

- **Independent variable** (what you change)
- **Dependent variable** (what you measure)
- **Control group** (for comparison)
- **Random assignment** (reduces bias)
]
]]

<div style="clear: both;"></div>

.font90[.center[*Example:* testing if **sunlight** affects **plant growth** by placing identical plants in different light conditions. &mdash; **Mostly quantitative**]]

---
### Quasi-Experiments: Real-World Testing

.pull-left[
```{r img_quasi, echo=FALSE, warning=FALSE, out.width="85%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-quasi.png")
grid.raster(img)
```
]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("book", fill = "#4a7fb5")` Definition**

Similar to experiments, **but without random assignment** or full control over all factors.
]

.content-box-green[
**`r fontawesome::fa("circle-question", fill = "#4a7fb5")` When to Use**

When true experiments aren't feasible or ethical, but you still want to test relationships between variables.
]

.content-box-yellow[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` Limitations**

Harder to establish causation &mdash; other factors might influence results. *Mostly quantitative.*
]
]]

---
### Surveys: Gathering Opinions and Data

.pull-left[
```{r img_survey, echo=FALSE, warning=FALSE, out.width="85%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-survey.png")
grid.raster(img)
```
]

.pull-right[.font90[
#### `r fontawesome::fa("tower-broadcast", height = "1em", fill = "#4a7fb5")` &nbsp; Broad Reach
Collect information from many people quickly.

#### `r fontawesome::fa("list-check", height = "1em", fill = "#4a7fb5")` &nbsp; Versatile Questions
Use multiple-choice, scales, or open-ended questions.

#### `r fontawesome::fa("chart-pie", height = "1em", fill = "#4a7fb5")` &nbsp; Quantifiable Results
Easily analyze and visualize responses.

#### `r fontawesome::fa("share-from-square", height = "1em", fill = "#4a7fb5")` &nbsp; Multiple Formats
Distribute on paper, online, or in person.
]]

.font80[.center[*Common in both quantitative and qualitative studies.*]]

---
### Interviews & Focus Groups: In-Depth Insights

.pull-left[
.content-box-blue[
**`r fontawesome::fa("comment", fill = "#4a7fb5")` Interviews**

One-on-one conversations to explore personal experiences and perspectives.

- Detailed individual responses
- Follow-up questions possible
- Good for sensitive topics
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("people-group", fill = "#4a7fb5")` Focus Groups**

Guided discussions with several participants to gather multiple viewpoints.

- Group interaction generates ideas
- See how opinions form and change
- Efficient for hearing from many people
]
]

<div style="clear: both;"></div>

.center[Both methods provide rich **qualitative** data that numbers alone can't capture.]

---
### Observation: Watching Real Behavior

.pull-left[
```{r img_obs, echo=FALSE, warning=FALSE, out.width="82%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-observation.png")
grid.raster(img)
```

.font80[.center[*Can be used in quantitative or qualitative studies.*]]

]

.pull-right[.font90[
#### `r fontawesome::fa("tree", height = "1em", fill = "#4a7fb5")` &nbsp; Natural Settings
Observe behavior as it naturally occurs, without intervention.

#### `r fontawesome::fa("list-check", height = "1em", fill = "#4a7fb5")` &nbsp; Structured or Unstructured
Use checklists for specific behaviors, or take open-ended notes.

#### `r fontawesome::fa("clock", height = "1em", fill = "#4a7fb5")` &nbsp; Time Sampling
Observe at different times to capture variation in behavior.

#### `r fontawesome::fa("user-secret", height = "1em", fill = "#4a7fb5")` &nbsp; Minimize Observer Effect
Be discreet so your presence doesn't change the behavior.
]]

---
### Case Studies: Deep Dives

.pull-left[
```{r img_case, echo=FALSE, warning=FALSE, out.width="85%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-casestudy.png")
grid.raster(img)
```
]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("book", fill = "#4a7fb5")` Definition**

Intensive study of a **single case** or small number of cases to gain in-depth understanding.
]

.content-box-green[
**`r fontawesome::fa("circle-question", fill = "#4a7fb5")` When to Use**

For unique situations, rare phenomena, or when you need comprehensive understanding of one example.
]

.content-box-yellow[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` Limitations**

Findings may not **generalize** to other cases or populations. *Mostly qualitative.*
]
]]

---
### Using Existing Data: Secondary Research

.pull-left[
```{r img_secondary, echo=FALSE, warning=FALSE, out.width="88%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-secondary.png")
grid.raster(img)
```
]

.pull-right[.font80[
#### `r fontawesome::fa("bolt", height = "1em", fill = "#4a7fb5")` &nbsp; Efficient Access
Use data already collected by others instead of gathering new information.

#### `r fontawesome::fa("clock-rotate-left", height = "1em", fill = "#4a7fb5")` &nbsp; Historical Perspective
Analyze trends over time periods you couldn't study directly.

#### `r fontawesome::fa("layer-group", height = "1em", fill = "#4a7fb5")` &nbsp; Multiple Sources
Government reports, research studies, archives, statistics, and historical documents.

#### `r fontawesome::fa("circle-check", height = "1em", fill = "#4a7fb5")` &nbsp; Source Verification
Always evaluate the credibility and quality of secondary data sources.
]]

.font80[.center[.content-box-yellow[**This is the method your camp project uses** &mdash; you'll analyze a confirmed, existing dataset.]]]

---
### Mixed Methods: Combining Approaches & Tools

.pull-left[
.content-box-blue[
**`r fontawesome::fa("puzzle-piece", fill = "#4a7fb5")` Benefits**

- Provides a **fuller picture** of the topic
- Combines breadth and depth
- Numbers + stories = comprehensive understanding
- Compensates for weaknesses in single methods
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("layer-group", fill = "#4a7fb5")` Examples**

- Survey + interviews
- Experiment + questionnaire
- Observation + focus groups
- Case study + statistical analysis
]
]

<div style="clear: both;"></div>

.center[Mixed methods require **careful planning** but provide rich, multidimensional insights.]

---
### Choosing the Right Method

.font90[
<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("circle-question", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Consider your research question</strong><br>
  What type of answer do you need? Numbers, descriptions, or both?</div>
</div>

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("box-open", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Assess your resources</strong><br>
  How much time, access, and materials do you have available?</div>
</div>

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("users", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Think about participants</strong><br>
  Who will you study, and how many can you realistically include?</div>
</div>

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("scale-balanced", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Weigh strengths and limitations</strong><br>
  Each method has trade-offs. Choose what best fits your specific needs.</div>
</div>
]

---
class: segue-yellow

# Activity: Design Your Study `r top_icon("drafting-compass")`

---
### Activity: Design Your Study

.pull-left[
.content-box-blue[
**`r fontawesome::fa("file-pen", fill = "#4a7fb5")` Complete the Session 5 Mission Sheet**

*"Research Design & Methodology"* &mdash; as a team, identify:

- `r fontawesome::fa("route", fill = "#4a7fb5")` Your **methodology**
- `r fontawesome::fa("database", fill = "#4a7fb5")` Your **dataset**
- `r fontawesome::fa("users", fill = "#4a7fb5")` Your **sample**
- `r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` Potential **limitations**
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("person-chalkboard", fill = "#4a7fb5")` Instructor circulates**

Your instructor will check the **feasibility** of each team's plan as you work.
]

<br>

.center[`r timer(25L)`]
]

---
### Scientific Storytelling: Methods Paragraph

.center[.font110[Write **one sentence** describing your methodology in plain language.]]

<br>

.content-box-blue[
.center[.font120[
*"This study uses **secondary data** from **[source]** to examine whether **[IV]** is associated with **[DV]** among **[population]**."*
]]
]

<br>

.pull-left[
.content-box-green[
**`r fontawesome::fa("comment-dots", fill = "#4a7fb5")` No jargon**

Plain language only &mdash; if a family member couldn't follow it, simplify.
]
]

.pull-right[
.content-box-yellow[
**`r fontawesome::fa("volume-high", fill = "#4a7fb5")` Read it aloud**

Share your sentence with the class.
]

.center[`r timer(10L)`]
]

---
### Tomorrow: We Continue Planning

.pull-left[
```{r img_plan2, echo=FALSE, warning=FALSE, out.width="88%", fig.show='hold', fig.align='center'}
img <- readPNG("assets/p5-plan.png")
grid.raster(img)
```
]

.pull-right[
<br>

## `r fontawesome::fa("bullseye", height = "1em", fill = "#4a7fb5")` &nbsp; Validity & Reliability

<br>

## `r fontawesome::fa("users", height = "1em", fill = "#4a7fb5")` &nbsp; Sampling

<br>

## `r fontawesome::fa("box-open", height = "1em", fill = "#4a7fb5")` &nbsp; Materials & Resources
]

---
class: segue

# Homework & Work Periods `r top_icon("pencil-alt")`

---
### Homework &mdash; Teamwork

.font80[
.pull-left[
.content-box-blue[
**`r fontawesome::fa("list-check", fill = "#4a7fb5")` Research Design Checklist**

Review the *"Research Design Checklist"* with your team members.
]

.content-box-blue[
**`r fontawesome::fa("file-lines", fill = "#4a7fb5")` Handout 4: Research Design Blueprint**

Start it and finish the **first half** of the items.
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("book-open", fill = "#4a7fb5")` Continue your lit review**

Review **1&ndash;2 more articles** and track them in the *"Source Tracker."* Identify **2&ndash;3 most important findings** and share your notes with a teammate.
]
]
]

---
### After-Class Team Session &mdash; 3:15&ndash;5:15 PM

.font80[*Teams work with TA / peer mentor available. Produce a tangible output before dinner. Not free time.*]

.font80[
.content-box-blue[
**`r fontawesome::fa("file-pen", fill = "#4a7fb5")` Complete the Research Design Mission Sheet** as a team &mdash; methodology, data source, sample description, potential limitations.
]

.content-box-blue[
**`r fontawesome::fa("database", fill = "#4a7fb5")` Confirm your dataset is finalized** &mdash; name, source, timeframe, row/column count. If collecting primary data, finalize your instrument today.
]

.content-box-green[
**`r fontawesome::fa("circle-check", fill = "#4a7fb5")` Due before dinner:** a completed Session 5 Mission Sheet &mdash; full methodology plan signed off by all members.
]
]

---
### Evening Work Period &mdash; 7:00&ndash;9:00 PM

.font80[*Individual + partner work during CTD RTA study hours. TA available for drop-in questions.*]

.font80[
.content-box-yellow[
**`r fontawesome::fa("pen", fill = "#4a7fb5")` Individual reflection**

What are your study's **two biggest limitations**? Write 3&ndash;4 sentences explaining each and how you might address them.
]

.content-box-blue[
**`r fontawesome::fa("database", fill = "#4a7fb5")` DataCORE Journal**

What is **one question about your data** you cannot yet answer? What analysis would you need to run?
]

.content-box-green[
**`r fontawesome::fa("file-lines", fill = "#4a7fb5")` (RcS) Draft the Methods section**

Past tense, 200&ndash;250 words, covering all fields from your Session 5 Mission Sheet.
]
]

---
class: segue-green

# Real-World PBL Case `r top_icon("balance-scale")`

---
### Economics PBL: Does Minimum Wage Affect Employment?

.font80[
.content-box-blue[
**`r fontawesome::fa("database", fill = "#4a7fb5")` The data**

**FRED dataset** of state-level **minimum wage** and **unemployment** data.
]

.pull-left[
.content-box-yellow[
**`r fontawesome::fa("circle-question", fill = "#4a7fb5")` Discuss as a team**

- How would you **design a study** to test whether minimum wage increases affect employment levels?
- What design would be **ideal**?
- What are the **limitations** of using this data?
- What **confounds** exist?
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` The big idea**

This case directly illustrates the difference between **observational** and **experimental** designs &mdash; and why we can rarely run the "ideal" experiment in the real world.
]
]
]

```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)
# https://hhadah.github.io/nexellence-summer-instit/week-01-foundations-of-inquiry/session-05-research-design-methodology/05-class.html

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-05-research-design-methodology/am/05-class.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-05-research-design-methodology/am/05-class.pdf")
to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-05-research-design-methodology/am/05-class.html")
```
