---
title: "DataCORE"
subtitle: "Portfolio Final Check, Data Art & Final Words"
date: "July 16, 2026"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | July 16, 2026</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "60%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
import seaborn as sns
sns.set_theme(style="whitegrid")
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.big-stat { font-size: 2.2rem; font-weight: 700; color: #2e8b57; }
.card-grid { display: flex; flex-wrap: wrap; gap: 0.7em; margin-top: 0.6em; }
.card { flex: 1 1 30%; background: #e7f3ec; border-radius: 8px; padding: 0.5em 0.8em; }
.card h4 { margin: 0 0 0.2em 0; color: #2e8b57; font-size: 0.95em; }
.card p { margin: 0; font-size: 0.8em; }
.chev-row-sm { display: flex; align-items: center; gap: 0.8em; margin-bottom: 0.3em; }
.chev-sm { flex: 0 0 50px; height: 56px; background: #cfe8d8;
  clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
  display: flex; align-items: center; justify-content: center;
  font-weight: bold; color: #1a4731; }
.chev-text { flex: 1; }
```


# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>DataCORE Portfolio Final Check <span style="color:#2e8b57;">(1:00&ndash;1:45)</span></h4></li>
    <li><h4>Data Art + Celebration Visualization <span style="color:#2e8b57;">(1:45&ndash;2:45)</span></h4></li>
    <li><h4>Scientific Storytelling &mdash; Final Words <span style="color:#2e8b57;">(2:45&ndash;3:00)</span></h4></li>
</ol>

.font90[*Lock the portfolio you'll display at the Summit, make one chart that's purely beautiful, and say out loud how far you've come.*]

```{python load-data, include=FALSE}
wages = pd.read_csv("wages.csv")
```

---
class: segue-green

# 1:00 &ndash; 1:45 &middot; DataCORE
## .white[Portfolio Final Check]

---
### `r fontawesome::fa("box-open", height = "1em", fill = "#2e8b57")` Six items, checked one last time

<div class="card-grid">
  <div class="card">
    <h4>`r fontawesome::fa("file-code", fill = "#2e8b57")` Analysis notebook</h4>
    <p>Clean, commented, readable by someone who did <em>not</em> write it.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("image", fill = "#2e8b57")` Exported figures</h4>
    <p>300 DPI, descriptive file names.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("table", fill = "#2e8b57")` Summary table</h4>
    <p>Statistics table with a proper caption.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("broom", fill = "#2e8b57")` Data cleaning log</h4>
    <p>What you changed, and why.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("robot", fill = "#2e8b57")` AI usage log</h4>
    <p>Tools, prompts, and errors the AI made.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("file-pdf", fill = "#2e8b57")` Notebook PDF</h4>
    <p>The exported document you display at the Summit.</p>
  </div>
</div>

.content-box-yellow[
.font85[`r fontawesome::fa("award", fill = "#c08a00")` &nbsp; **This portfolio is a real artifact.** Future versions of Nexellence INQUIRE may cite it as an example &mdash; treat it like a first publication.]
]

---
### `r fontawesome::fa("magnifying-glass", height = "1em", fill = "#2e8b57")` The "someone else" test

.pull-left[
.font90[
Hand your notebook to a person who has never seen it. Can they, **without you in the room**:

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("play", fill = "#2e8b57")`</div><div class="chev-text"><strong>Run it</strong> top to bottom with no errors?</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("book-open", fill = "#2e8b57")`</div><div class="chev-text"><strong>Follow it</strong> from the header cell to the limitations cell?</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("chart-simple", fill = "#2e8b57")`</div><div class="chev-text"><strong>Rebuild your main figure</strong> from your code?</div></div>
]
]

.pull-right[
.content-box-green[
.font85[`r fontawesome::fa("circle-check", fill = "#009E73")` &nbsp; If the answer to all three is **yes**, your work is reproducible &mdash; the highest compliment a portfolio can earn.]
]
]

.center[`r timer(45L)`]

---
class: segue-green

# 1:45 &ndash; 2:45 &middot; Data Art
## .white[Celebration Visualization]

---
### `r fontawesome::fa("palette", height = "1em", fill = "#2e8b57")` Make one chart that's beautiful *and* true

.pull-left[
.font90[
Create one **"data art"** visualization from your team's data &mdash; beautiful and informative at once. **No rules on format:**

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("cloud", fill = "#2e8b57")`</div><div class="chev-text">Word cloud of key variables</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("film", fill = "#2e8b57")`</div><div class="chev-text">Animated plot</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("chart-area", fill = "#2e8b57")`</div><div class="chev-text">An unusual type: ridgeline, waffle, network graph</div></div>
]
]

.pull-right[
```{python dataart, echo=FALSE, out.width="99%", fig.width=5.4, fig.height=3.7}
import seaborn as sns
order = wages.groupby("education")["hourly_wage"].median().sort_values().index
plt.figure()
sns.violinplot(data=wages, x="hourly_wage", y="education", order=order,
               palette="viridis", cut=0, linewidth=0.8, inner="quart")
plt.title("The shape of wages, by education", fontsize=13, fontweight="bold")
plt.xlabel("Hourly wage ($)", fontsize=12); plt.ylabel("")
plt.tight_layout(); plt.show()
```
.font70[.center[*One example &mdash; a violin "ridgeline" in viridis.*]]
]

---
### `r fontawesome::fa("share-nodes", height = "1em", fill = "#2e8b57")` Share &amp; celebrate

.pull-left[.font90[
An informal **show-and-tell**:

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("eye", fill = "#2e8b57")`</div><div class="chev-text">Put your data art on the screen.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("comment", fill = "#2e8b57")`</div><div class="chev-text">Say one sentence about what makes it <em>you</em>.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("hands-clapping", fill = "#2e8b57")`</div><div class="chev-text">Applaud every team.</div></div>
]]

.pull-right[.font90[
.content-box-yellow[
**`r fontawesome::fa("champagne-glasses", fill = "#c08a00")` Three weeks, one artifact**

You turned a question into data, data into findings, and findings into a story. That's the whole of research &mdash; take a moment to see it.
]
]]

.center[`r timer(60L)`]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[Final Words]

---
### `r fontawesome::fa("quote-left", height = "1em", fill = "#2e8b57")` One sentence each

.font90[Every team member writes and shares a single sentence:]

.content-box-blue[
.center[.font110[*"Three weeks ago I didn't know how to __________.*

*Now I can __________."*]]
]

.pull-left[
.content-box-green[
.font85[`r fontawesome::fa("microphone", fill = "#009E73")` &nbsp; Say it out loud to the group. This is the **closing statement of your research apprenticeship**.]
]
]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("door-open", fill = "#c08a00")` &nbsp; &hellip;and the **opening statement** of whatever comes next.]
]
]

.center[`r timer(15L)`]

---
class: segue-green

# .white[You came in curious. You're leaving a researcher.]
## .white[Congratulations &mdash; and see you at the Summit.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-14-final-preparations-reflection/pm/14-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-14-final-preparations-reflection/pm/14-pm-datacore.pdf")
```
