---
title: "Research Methods"
subtitle: "Day 14 · Session 14: Final Preparations, Reflection & Camp Review"
# author: "Hussain Hadah (he/him)"
date: "July 16, 2026"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | July 16, 2026</div>

<!--- `r rmarkdown::metadata$subtitle` | July 16, 2026-->

---

class: title-slide
background-position: 10% 90%, 100% 50%
background-size: 160px, 50% 100%
background-color: #0148A4

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

    .img-row {
      display: flex;
      gap: 1em;
      align-items: center;
      justify-content: center;
      margin-top: 0.8em;
    }
    .img-row img {
      max-width: 100%;
      border-radius: 6px;
    }
    .img-row .cell {
      flex: 1;
      text-align: center;
    }

    .remark-code {
      font-size: 0.75em !important;
    }

    .card-grid {
      display: flex;
      flex-wrap: wrap;
      gap: 0.8em;
      margin-top: 0.6em;
    }

```

# .text-shadow[.white[Outline for Today]]

<ol>
    <li><h4 class="white">Final Polish Time</h4></li>
    <li><h4 class="white">Camp Review & Synthesis Discussion</h4></li>
    <li><h4 class="white">Camp Review Mission Sheet</h4></li>
    <li><h4 class="white">Letter to Future Nexellence INQUIRE Students</h4></li>
    <li><h4 class="white">Scientific Storytelling: Opening Line Practice</h4></li>
</ol>

---
class: segue-yellow

# Session 14: Final Preparations, <br> Reflection & Camp Review `r top_icon("flag-checkered")`

---
### Learning Objectives

.font90[
#### `r fontawesome::fa("wand-magic-sparkles", height = "1em", fill = "#4a7fb5")` &nbsp; Polish presentations to final presentation quality
Last changes, final checks &mdash; everything Summit-ready.

#### `r fontawesome::fa("brain", height = "1em", fill = "#4a7fb5")` &nbsp; Reflect on learning and growth across three weeks
Name what you know now that you didn't when you started.

#### `r fontawesome::fa("trophy", height = "1em", fill = "#4a7fb5")` &nbsp; Understand and celebrate the full research process
See the whole arc &mdash; question to Summit &mdash; that you completed.

#### `r fontawesome::fa("people-group", height = "1em", fill = "#4a7fb5")` &nbsp; Prepare for the Research Summit audience
Parents, guests, and instructors &mdash; ready your hook and your poise.
]

---
class: segue-yellow

# Morning: ResearchCORE `r top_icon("sun")`

---
class: segue-green

# Final Polish Time `r top_icon("magic")`

---
### Final Polish Time (45 min)

.pull-left[.font90[
Make any **remaining changes** to your slides &mdash; then stop tinkering.

While you polish, the instructor does a **final review** of each team's:

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("quote-right", fill = "#0148A4")`</div><div class="chev-text"><strong>References slide</strong> &mdash; every cited source listed.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("robot", fill = "#0148A4")`</div><div class="chev-text"><strong>AI Attribution Statement</strong> &mdash; present and accurate.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("equals", fill = "#0148A4")`</div><div class="chev-text"><strong>Citation format</strong> &mdash; consistent throughout.</div></div>
]]

.pull-right[.font90[
.content-box-yellow[
**`r fontawesome::fa("hand", fill = "#c08a00")` Know when to stop**

At some point, more edits add risk, not polish. When the checklist is green, close the laptop and rehearse instead.
]
]]

.center[`r timer(45L)`]

---
class: segue-green

# Camp Review & <br> Synthesis Discussion `r top_icon("comments")`

---
### Camp Review & Synthesis (30 min)

.pull-left[.font90[
An instructor-led discussion. One prompt:

.content-box-blue[
**What are the five most important things you know about research that you *didn't* know three weeks ago?**
]

Each student contributes **one idea.** We capture them all on the whiteboard under one heading:

.content-box-green[
.center[**"What Nexellence INQUIRE taught us"**]
]
]]

.pull-right[.font90[
.content-box-yellow[
**`r fontawesome::fa("lightbulb", fill = "#c08a00")` Think across the whole arc**

From a first curious question, to methods, to data, to a story you can defend &mdash; what shifted in how you *think*, not just what you can do?
]
]]

.center[`r timer(30L)`]

---
### Session 14 Mission Sheet: Camp Review (30 min)

.font90[Complete four reflections &mdash; your record of the journey:]

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>One thing that went well</strong> in your rehearsal.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">Something <strong>interesting you learned from another team</strong>.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">What you're <strong>most proud of</strong>.</div></div>
<div class="chev-row-sm"><div class="chev-sm">4</div><div class="chev-text">A <strong>new research question</strong> you'd love to explore in the future.</div></div>

.content-box-blue[
.font85[`r fontawesome::fa("seedling", fill = "#4a7fb5")` &nbsp; That last one matters most: the best sign of a real researcher is leaving with **more** questions than you arrived with.]
]

.center[`r timer(30L)`]

---
class: segue-yellow

# Letter to Future <br> Nexellence INQUIRE Students `r top_icon("envelope")`

---
### Letter to a Future Student (30 min)

.pull-left[.font90[
Write a **one-page letter** &mdash; or record a voice note or video &mdash; to a future Nexellence INQUIRE student.

**Speak to them honestly:**

- What **advice** would you give?
- What **surprised** you along the way?
- What would you tell your **past self** before starting?
]]

.pull-right[.font90[
.content-box-green[
**`r fontawesome::fa("box-archive", fill = "#009E73")` Your words will outlast you here**

These letters are archived &mdash; future students really will read them. Write the thing you *wish* someone had told you on day one.
]
]]

.center[`r timer(30L)`]

---
class: segue-yellow

# Scientific Storytelling: <br> Opening Line Practice `r top_icon("bullhorn")`

---
### Your First 30 Seconds

.pull-left[.font90[
Each presenter rehearses their **opening line** and first-slide intro **one final time.**

.content-box-yellow[
**`r fontawesome::fa("ban", fill = "#c08a00")` The one rule:** no one starts with *"Hi, my name is&hellip;"*
]

Open with a **hook** instead:
- a **surprising statistic**
- a **compelling question**
- a **striking image**
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("wave-square", fill = "#4a7fb5")` The first 30 seconds sets the tone**

An audience decides whether to lean in almost immediately. Earn their attention before you spend it on introductions.
]

.content-box-green[
.font88[*"Every year, teenagers lose two full weeks of sleep to their phones. We wanted to know what that costs them&hellip;"*]
]
]]

.center[`r timer(15L)`]

---
### Session 14: What You Can Do Now

.font80[
**Final polish:**

- Bring your slides, **References**, and **AI attribution** to final, consistent, Summit quality &mdash; and recognize when to stop editing and start rehearsing.

**Reflection & growth:**

- Articulate the **five biggest things** you learned about research, complete your Camp Review reflections, and write a letter that passes what you learned to the next cohort.

**Summit readiness:**

- Open with a **hook**, not "Hi, my name is" &mdash; a stat, a question, or an image that earns the room's attention in the first 30 seconds.
]
