---
title: "Research Methods"
subtitle: "Day 11 · Session 11: Interpreting Findings & Discussion"
# author: "Hussain Hadah (he/him)"
date: "July 13, 2026"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | July 13, 2026</div>

<!--- `r rmarkdown::metadata$subtitle` | July 13, 2026-->

---

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
timer_sm <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0 0 0 0.3em; vertical-align: middle;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "0.9rem",
    padding = "3px 6px",
    margin = "0",
    warn_when = 30L,
    ...
  )
}
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

    .img-row {
      display: flex;
      gap: 1em;
      align-items: center;
      justify-content: center;
      margin-top: 0.8em;
    }
    .img-row img {
      max-width: 100%;
      border-radius: 6px;
    }
    .img-row .cell {
      flex: 1;
      text-align: center;
    }

    .remark-code {
      font-size: 0.75em !important;
    }

```

### Warm-Up &mdash; Show Us Your Figures (20 min)

.pull-left[.font85[
Each team takes the floor for **~3 minutes** and presents **2&ndash;3 figures** from their results.

#### `r fontawesome::fa("chart-line", height = "1em", fill = "#4a7fb5")` &nbsp; What to show
Your **best** figures &mdash; the ones you would actually put in the paper or the talk.

#### `r fontawesome::fa("comment", height = "1em", fill = "#4a7fb5")` &nbsp; How to present each one
One sentence on **what the figure shows**, one on **what it means**. No table-speak.

#### `r fontawesome::fa("users", height = "1em", fill = "#4a7fb5")` &nbsp; Audience job
Say one thing that **works** and one thing that would make the figure **clearer**.
]]

.pull-right[
.content-box-blue[
.font85[
**For each figure, be ready to answer:**

- What is on the **x- and y-axis**, and what is the unit?
- What is the **one pattern** you want us to see?
- Does it stand on its own &mdash; **title, labels, legend**, readable text?
- Which figure is your **headline** result?
]
]

.content-box-yellow[
.font85[**Keep these figures handy** &mdash; you'll interpret them in the Discussion we build today.]
]
.center[`r timer(20L)`]
]


---
### Learning Objectives

.font90[
#### `r fontawesome::fa("diagram-project", height = "1em", fill = "#4a7fb5")` &nbsp; Connect findings to the question &amp; literature
Weave your data findings into a coherent narrative that ties back to the research question and to what others have found.

#### `r fontawesome::fa("scale-balanced", height = "1em", fill = "#4a7fb5")` &nbsp; Write a Discussion that contextualizes without overclaiming
Interpret your results honestly &mdash; say what they mean, and no more than the data can support.

#### `r fontawesome::fa("triangle-exclamation", height = "1em", fill = "#4a7fb5")` &nbsp; Identify limitations honestly and constructively
Name the caveats that most threaten your interpretation &mdash; and what they imply.

#### `r fontawesome::fa("layer-group", height = "1em", fill = "#4a7fb5")` &nbsp; Begin integrating all sections into a complete presentation
Assemble intro, methods, results, discussion, and conclusion into one connected story.
]

---
### From Results to Meaning

.pull-left[.font90[
.content-box-yellow[
**Results tell the audience *what* you found. The Discussion tells them *what it means*.**
]

The Results section reports the numbers. The Discussion is where you become the interpreter of your own evidence.

**A strong Discussion answers four questions:**

- **So what?** &mdash; why should anyone care?
- **Why?** &mdash; what might explain the finding?
- **Compared to what?** &mdash; how does it sit against the literature?
- **What are the caveats?** &mdash; where could you be wrong?
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` The interpreter's job**

You are the person who knows this dataset best. The Discussion is your chance to explain what the reader should take away &mdash; and to be honest about the limits of that claim.
]

.content-box-green[
**`r fontawesome::fa("comments", fill = "#4a7fb5")` Discussion &ne; repeat of Results**

Don't just restate every number again. Pick the findings that matter and interpret them.
]
]]

---
class: segue-green

# How to Write a <br> Discussion Section `r top_icon("pen-fancy")`

---
### The Four-Move Structure

.font85[Every paragraph of a good Discussion moves through the same four steps:]

<div class="chev-flow">
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("comment", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>1. Restate</strong><br>
    Say the finding in plain language &mdash; no jargon, no table numbers.</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("gears", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>2. Explain</strong><br>
    Why might this be happening? What is the mechanism?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("scale-balanced", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>3. Compare</strong><br>
    Agree, disagree, or add nuance to what the literature said?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("triangle-exclamation", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>4. Limit</strong><br>
    Name the limitation that most threatens the interpretation.</p>
  </div>
</div>

--

.content-box-yellow[
.font85[**Use it as a template.** Take one key finding, walk it through all four moves, and you have a complete Discussion paragraph.]
]

---
### Move 1 &mdash; Restate the Finding in Plain Language

.pull-left[.font85[
.content-box-blue[
**`r fontawesome::fa("comment", fill = "#4a7fb5")` Plain language, not table-speak**

Translate the statistic into a sentence a non-expert would understand.
]

- Drop the *p*-values and coefficients for a moment.
- Say what actually happened to whom.
- Anchor it to your research question.
]]

.pull-right[.font85[
.content-box-yellow[
**`r fontawesome::fa("xmark", fill = "#D55E00")` Instead of:**

*"The coefficient on treatment was 4.2 (p &lt; 0.05)."*
]

.content-box-green[
**`r fontawesome::fa("check", fill = "#009E73")` Say:**

*"Students who used the study app scored about 4 points higher on the final test than those who did not."*
]
]]

---
### Move 2 &mdash; Explain It

.pull-left[.font85[
**Why might you have found this?** Offer a *mechanism* &mdash; a plausible story for how the cause produces the effect.

- What do we already know about *why* this relationship could exist?
- Is there a chain of steps linking your variables?
- Could something else be driving it? (name the **alternative explanation**)
]]

.pull-right[.font85[
.content-box-blue[
**`r fontawesome::fa("gears", fill = "#4a7fb5")` Example mechanism**

*"The app may help because it spaces practice over time, and spaced repetition is known to improve retention."*
]

.content-box-yellow[
**`r fontawesome::fa("circle-question", fill = "#4a7fb5")` Alternative explanation**

*"But motivated students may have both chosen the app and studied more &mdash; so we can't rule out self-selection."*
]
]]

---
### Move 3 &mdash; Compare to the Literature

.font85[Does your finding **agree**, **disagree**, or **add nuance** to what was already known?]

.three-col[
.col[
`r fontawesome::fa("thumbs-up", height = "2.2em", fill = "#009E73")`

**Agrees**

*"This is consistent with Smith (2021), who also found a positive effect of spaced practice."*
]
.col[
`r fontawesome::fa("code-compare", height = "2.2em", fill = "#D55E00")`

**Disagrees**

*"In contrast to Lee (2020), we found no gender difference &mdash; possibly because our sample was older."*
]
.col[
`r fontawesome::fa("plus", height = "2.2em", fill = "#4a7fb5")`

**Adds nuance**

*"Building on Gomez (2022), we show the effect holds, but only for students who used the app weekly."*
]
]

--

.content-box-blue[
.font85[**Every comparison needs a citation.** Tie each finding to at least one source from your literature review.]
]

---
### Move 4 &mdash; Identify the Key Limitation

.pull-left[.font85[
Every study has limits. **Name the one that most threatens your interpretation** &mdash; not a laundry list.

**Ask:**
- What would a skeptic attack first?
- Does this limitation change *what the finding means*?
- What would you do differently with more time or data?
]]

.pull-right[.font85[
.content-box-yellow[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` A limitation with teeth**

*"Because students chose whether to use the app, our estimate may reflect who they are, not what the app does. A randomized trial would settle this."*
]

.content-box-green[
**`r fontawesome::fa("check", fill = "#009E73")` Honest &ne; apologetic**

State the limit, say what it implies, and move on. Don't undermine your whole study.
]
]]

---
### Interpret Without Overclaiming

.font80[
.pull-left[
.content-box-yellow[
**`r fontawesome::fa("xmark", fill = "#D55E00")` Overclaiming**

- *"This proves the app causes higher scores."*
- *"Our results show this will work everywhere."*
- Treating correlation as causation.
- Generalizing beyond your sample.
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("check", fill = "#009E73")` Calibrated claims**

- *"Our results **suggest** the app **is associated with** higher scores."*
- *"**Within our sample**, students who&hellip;"*
- Hedging language: *may, suggests, is consistent with*.
- Scope conditions: for whom, where, when.
]
]
]

.content-box-blue[
.font85[**The test:** could you defend every sentence of your Discussion if the author you cited were sitting in the front row?]
]

---
class: segue-yellow

# Synthesizing Your Findings <br> with the Literature `r top_icon("book-open")`

---
### What Does Synthesis Mean?

.pull-left[.font90[
.content-box-yellow[
**Synthesis is more than summary.** It means combining ideas from different sources to see *how they relate* &mdash; not just what each one says.
]

A good discussion doesn't list studies one after another. It **weaves** them together around *your* finding.

**Ask of your sources:**
- Do the authors **agree** or **disagree**?
- Are they studying the same issue from **different angles**?
- Does one source **support, expand on, or challenge** another?
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("diagram-project", fill = "#4a7fb5")` The "discuss" checklist**

To *discuss* a result means to:

1. Restate the key finding
2. Compare to what the literature said
3. Offer an explanation for it
4. Acknowledge limitations
5. Suggest what comes next
]
]]

---
### Group Ideas, Not Authors

.pull-left[.font85[
.content-box-yellow[
**`r fontawesome::fa("xmark", fill = "#D55E00")` Don't organize by author:**

*"Smith says&hellip; then Jones says&hellip; then Lee says&hellip;"*
]

This reads like a list, not an argument.
]]

.pull-right[.font85[
.content-box-green[
**`r fontawesome::fa("check", fill = "#009E73")` Organize by theme, issue, or trend:**

*"Many studies agree that social media affects attention span (Smith, 2021; Lee, 2020), but a few challenge this idea (Gomez, 2022)."*
]

One sentence, three sources, one clear point.
]]

---
### Use Comparative Language

.pull-left[.font90[
.content-box-blue[
**`r fontawesome::fa("link", fill = "#4a7fb5")` Helpful connecting phrases**

- *"Similarly&hellip;"*
- *"In contrast&hellip;"*
- *"Building on&hellip;"*
- *"While some researchers argue&hellip;, others suggest&hellip;"*
- *"There is a general consensus that&hellip;"*
]
]]

.pull-right[.font90[
.content-box-green[
**`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` Highlight gaps or conflicts**

Point out where sources disagree or where more research is needed.

*"While most studies link sleep to better memory, few explore how this differs by age group."*
]
]]

--

.content-box-yellow[
.font85[**End with the big picture:** what have you learned from all the sources *combined*? What patterns, debates, or gaps set the stage for your own finding?]
]

---
class: segue-green

# Presentations: Discussion <br> + Conclusion Slides `r top_icon("chalkboard")`

---
### What the Ending Has to Do

.pull-left[.font90[
The last two slides of your talk carry the whole argument home.

**Build your Discussion + Conclusion slides around three questions:**

- What do our results **mean**?
- Did they **support or challenge** our hypothesis?
- **Who should care** about this?
]]

.pull-right[.font90[
.content-box-blue[
**`r fontawesome::fa("arrows-left-right", fill = "#4a7fb5")` Does the discussion connect back?**

The findings must tie clearly to the **original research question**.
]

.content-box-green[
**`r fontawesome::fa("flag-checkered", fill = "#4a7fb5")` Does the conclusion land?**

The audience should leave with **one clear takeaway**.
]
]]

---
### Anatomy of a Strong Conclusion Slide

.font85[
.pull-left[
.content-box-green[
**`r fontawesome::fa("bullseye", fill = "#4a7fb5")` One takeaway**

Lead with the single sentence you want the audience to remember.
]

.content-box-green[
**`r fontawesome::fa("link", fill = "#4a7fb5")` Answers the question**

Explicitly close the loop back to the question you opened with.
]
]

.pull-right[
.content-box-blue[
**`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` Honest caveat**

One limitation, stated plainly &mdash; it builds credibility.
]

.content-box-blue[
**`r fontawesome::fa("arrow-right", fill = "#4a7fb5")` Where next**

A forward-looking line: what this opens up or what should follow.
]
]
]

.content-box-yellow[
.font85[**Avoid** ending on a wall of numbers or a plain "Thank you." End on *meaning*.]
]

---
class: segue-yellow

# Scientific Storytelling: <br> Presentation Transitions `r top_icon("comments")`

---
### Verbal Transitions Between Sections

.font90[Smooth transitions signal a **well-told story**. These are the sentences that carry your audience from one section to the next.]

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("arrow-right", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Methods &rarr; Results:</strong> <em>"Now that we've seen the methods, let me walk you through what we found&hellip;"</em></div>
</div>

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("arrow-right", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Results &rarr; Discussion:</strong> <em>"Our results suggest that&hellip; This aligns with / differs from what [Author, Year] found in their study&hellip;"</em></div>
</div>

<div class="chev-row">
  <div class="chev-sm">`r fontawesome::fa("arrow-right", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Discussion &rarr; Conclusion:</strong> <em>"So what does all of this mean? Here is the one thing we want you to remember&hellip;"</em></div>
</div>

--

.content-box-blue[
.font85[**Practice out loud.** The transition is where a presentation feels either seamless or stitched together. Rehearse the hand-offs, especially when the speaker changes &mdash; you'll do this in the activity block.]
]

---
### Putting the Whole Story Together

.font85[
<div class="chev-flow">
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("circle-question", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Question</strong><br>
    What did we set out to answer?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("flask", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Methods</strong><br>
    How did we go about answering it?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("chart-line", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Results</strong><br>
    What did we find?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("lightbulb", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Discussion</strong><br>
    What does it mean, and how does it fit?</p>
  </div>
  <div class="chev-step">
    <div class="chev-right">`r fontawesome::fa("flag-checkered", height = "1.6em", fill = "#0148A4")`</div>
    <p><strong>Conclusion</strong><br>
    So what &mdash; and what next?</p>
  </div>
</div>
]

.content-box-yellow[
.font85[**Integration is today's goal:** every section should point back to the question and forward to the takeaway. That through-line *is* the story.]
]

---
class: segue-green

# In-Class Activities `r top_icon("users")`

---
### In-Class Activities &mdash; Run in This Order (90 min)

.font70[
.pull-left[
.content-box-blue[
**1. `r fontawesome::fa("file-pen", fill = "#4a7fb5")` Draft Report Feedback** &mdash; 20 min &nbsp; `r timer_sm(20L)`

Marked-up drafts returned; two class-wide teaching points. Teams read the comments together and ask: *which fix would improve the paper most? What can we fix today?*
]

.content-box-yellow[
**2. `r fontawesome::fa("pen-to-square", fill = "#4a7fb5")` Team Activity: Discussion Synthesis** &mdash; 30 min &nbsp; `r timer_sm(30L)`

Write one paragraph on your **main finding** using the four moves (restate &rarr; explain &rarr; compare &rarr; limit), citing **at least one source**. Expect probing questions: *Why do you think you found that? What alternative explanation is there? Who would challenge this?*

**Mission sheet &mdash; four lines:** (1) main conclusion in plain language; (2) comparison to literature, with citation; (3) one important limitation; (4) one next step. Then swap with a neighboring team for quick feedback.
]
]

.pull-right[
.content-box-green[
**3. `r fontawesome::fa("chalkboard", fill = "#4a7fb5")` Present the Ending** &mdash; 30 min &nbsp; `r timer_sm(30L)`

Each team presents their **Discussion + Conclusion slides**. Feedback focuses on two questions:

- Does the discussion **connect back** to the original research question?
- Does the conclusion **land** on one clear takeaway?
]

.content-box-blue[
**4. `r fontawesome::fa("comments", fill = "#4a7fb5")` Transition Practice** &mdash; 10 min &nbsp; `r timer_sm(10L)`

Rehearse the hand-offs **out loud**: Methods &rarr; Results &rarr; Discussion &rarr; Conclusion &mdash; especially where the speaker changes.
]
]
]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)
# https://hhadah.github.io/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-11-interpreting-findings-discussion/am/11-class.html

to_pdf(from = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-11-interpreting-findings-discussion/am/11-class.html",
       to = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-11-interpreting-findings-discussion/am/11-class.pdf")
to_pdf(from = "~/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-11-interpreting-findings-discussion/am/11-class.html")
```
