---
title: "Research Methods"
subtitle: "Day 10 · Synthesizing Findings & Polishing the Middle"
# author: "Hussain Hadah (he/him)"
date: "July 10, 2026"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | July 10, 2026</div>

<!--- `r rmarkdown::metadata$subtitle` | July 10, 2026-->

---

class: title-slide
background-position: 10% 90%, 100% 50%
background-size: 160px, 50% 100%
background-color: #0148A4

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

    .img-row {
      display: flex;
      gap: 1em;
      align-items: center;
      justify-content: center;
      margin-top: 0.8em;
    }
    .img-row img {
      max-width: 100%;
      border-radius: 6px;
    }
    .img-row .cell {
      flex: 1;
      text-align: center;
    }

    .remark-code {
      font-size: 0.75em !important;
    }

```


```{css, echo=F}
    /* 2x2 card grid used for "Characteristics" and "Tips" slides */
    .card-grid {
      display: flex;
      flex-wrap: wrap;
      gap: 0.8em;
      margin-top: 0.8em;
    }
    .card {
      flex: 1 1 44%;
      background: #d9ecf9;
      border-radius: 8px;
      padding: 0.5em 0.9em;
    }
    .card h4 {
      margin: 0 0 0.25em 0;
      color: #0148A4;
    }
    .card p {
      margin: 0;
      font-size: 0.85em;
    }
```

```{r countdown-setup, echo=FALSE}
library(countdown)
# Helper: a centered, inline countdown timer (click to start/pause/reset)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.6em auto;",
    color_border = "#0148A4",
    color_running_background = "#0148A4",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2.2rem",
    warn_when = 30L,
    ...
  )
}
```


# .text-shadow[.white[Outline for Today]]

<ol>
    <li><h4 class="white">Methods + Results Presentations: "The Middle"</h4></li>
    <li><h4 class="white">Results vs. Discussion vs. Conclusion</h4></li>
    <li><h4 class="white">Synthesizing Results with the Literature</h4></li>
    <li><h4 class="white">Synthesizing Findings Mission Sheet</h4></li>
    <li><h4 class="white">Scientific Storytelling: Discussion Slides Draft</h4></li>
</ol>

---
### Welcome to Day 10!

.pull-left[.font90[
Today you **synthesize** &mdash; turning a pile of results into a coherent story, and connecting it back to the literature.

By the end of the day you will be able to:

- **Synthesize** your results with existing literature into a coherent interpretation
- **Differentiate** results from discussion and conclusion
- **Strengthen** your methods and results sections
- Understand the role of **limitations** in scientific humility
]]

.pull-right[
.content-box-blue[
.font90[`r fontawesome::fa("layer-group", fill = "#4a7fb5")` &nbsp; **"The Middle" of your paper** is where data becomes meaning.
]
]
.content-box-green[
.font90[`r fontawesome::fa("book-open", fill = "#009E73")` &nbsp; A finding only matters once you say **what it means** and **how it fits** what others have found.
]
]
]

---
class: segue-yellow

# Methods + Results Presentations: <br> "The Middle" `r top_icon("microphone")`

---
### Methods + Results Presentations `r fontawesome::fa("clock", fill="#fff", height="0.8em")` 40 min

.content-box-blue[
Each team presents its **Methods** and **Results** slides &mdash; same structured format: **2 min present + 2 min feedback.**
]

.pull-left[.font90[
**As the audience, focus on three questions:**

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text">Are the <strong>results clear</strong>?</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">Are the <strong>figures interpretable</strong>?</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">Is the <strong>methods description accurate</strong>?</div></div>
]]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("comments", fill = "#c08a00")` &nbsp; Feedback names **one strength** and **one specific fix** &mdash; "the axis needs units," not "nice chart."]
]

.center[`r timer(40L)`]
]

---
class: segue-green

# Results vs. Discussion <br> vs. Conclusion `r top_icon("layer-group")`

---
### Three Different Jobs

.font90[The biggest write-up confusion: blurring three sections that each do **one** job.]

<div class="three-col">
  <div class="col">
    <div class="content-box-blue">
    <p><strong>`r fontawesome::fa("magnifying-glass-chart", fill="#4a7fb5")` Results</strong></p>
    <p class="font80"><strong>What you found.</strong> Facts and numbers, no opinion. <em>"Group A scored 8 points higher (p = 0.01)."</em></p>
    </div>
  </div>
  <div class="col">
    <div class="content-box-green">
    <p><strong>`r fontawesome::fa("lightbulb", fill="#009E73")` Discussion</strong></p>
    <p class="font80"><strong>What it means.</strong> Interpret, compare to the literature, explain, note limits. <em>"This aligns with Lee (2020)…"</em></p>
    </div>
  </div>
  <div class="col">
    <div class="content-box-yellow">
    <p><strong>`r fontawesome::fa("flag-checkered", fill="#c08a00")` Conclusion</strong></p>
    <p class="font80"><strong>So what?</strong> The big-picture takeaway and why it matters. Short, no new data.</p>
    </div>
  </div>
</div>

--

.content-box-blue[
.font90[**Quick test:** a sentence with a **number** is usually a *result*; a sentence with **"this suggests / because / compared to"** is usually *discussion*.]
]

---
### Spot the Section

.content-box-blue[
Which section does each sentence belong in &mdash; **Results**, **Discussion**, or **Conclusion**?
]

.font90[
1. *"Students who slept more scored 6 points higher on average."*
2. *"This supports Garcia (2019), who linked sleep to memory consolidation."*
3. *"Schools should consider later start times to support learning."*
]

--

.font90[
1. `r fontawesome::fa("magnifying-glass-chart", fill = "#4a7fb5")` **Results** &mdash; a measured fact.
2. `r fontawesome::fa("lightbulb", fill = "#009E73")` **Discussion** &mdash; compares to the literature.
3. `r fontawesome::fa("flag-checkered", fill = "#c08a00")` **Conclusion** &mdash; the broader "so what."
]

---
class: segue-yellow

# Synthesizing Results <br> with the Literature `r top_icon("book-open")`

---
### What Does "Discuss" Actually Mean? `r fontawesome::fa("clock", fill="#fff", height="0.8em")` 25 min

.font90[To **discuss** your findings is to do **five things, in order**:]

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>Restate</strong> your key finding (in plain language).</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text"><strong>Compare</strong> it to what the literature said &mdash; agree? disagree? extend?</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text"><strong>Explain</strong> <em>why</em> you might have found what you found.</div></div>
<div class="chev-row-sm"><div class="chev-sm">4</div><div class="chev-text"><strong>Acknowledge</strong> your limitations.</div></div>
<div class="chev-row-sm"><div class="chev-sm">5</div><div class="chev-text"><strong>Suggest</strong> what comes next.</div></div>

--

.content-box-green[
.font90[This is the skeleton of **every** Discussion section you'll ever read. Master the order and the writing gets much easier.]
]

---
### Synthesis in Action

.pull-left[.font82[
**A worked Discussion mini-paragraph:**

.content-box-blue[
*"We found that students with more screen time slept about 1.3 hours less **(restate)**. This echoes Twenge (2017), who reported the same link in a national sample **(compare)**. A likely mechanism is delayed bedtime from late-night use **(explain)**. Our sample was small and self-reported **(limitation)**, so a larger study with objective sleep tracking is a natural next step **(next)**."*
]
]]

.pull-right[.font85[
.content-box-yellow[
**`r fontawesome::fa("link", fill = "#c08a00")` Every finding earns a citation.**

Each result in your Discussion should connect to **at least one source** from your literature review &mdash; agreement *or* tension both count.
]
]]

---
### Limitations: Scientific Humility

.pull-left[.font88[
Naming limitations isn't weakness &mdash; it's **credibility.** It shows you understand what your study can and cannot claim.

.content-box-blue[
**Common, honest limitations:**
- Small or non-random **sample**
- **Self-reported** or noisy measures
- **Observational** data &rarr; correlation, not proof of cause
- Short **time window**
]
]]

.pull-right[.font88[
.content-box-green[
**`r fontawesome::fa("scale-balanced", fill = "#009E73")` Frame each limitation forward:**

*"Because our sample was one school, results may not generalize &mdash; a multi-school study would test this."*
]

.content-box-yellow[
.font90[`r fontawesome::fa("triangle-exclamation", fill = "#c08a00")` &nbsp; A limitation **bounds** your claim; it doesn't erase it.]
]
]]

---
class: segue-green

# Synthesizing Findings <br> Mission Sheet `r top_icon("clipboard")`

---
### Session 10 Mission Sheet `r fontawesome::fa("clock", fill="#fff", height="0.8em")` 35 min

.content-box-blue[
As a team, complete the **"Synthesizing Findings"** Mission Sheet &mdash; four sentences.
]

.font88[
<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>Main conclusion</strong> &mdash; your single biggest takeaway.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text"><strong>Comparison to literature</strong> &mdash; one source you agree or disagree with.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text"><strong>One important limitation.</strong></div></div>
<div class="chev-row-sm"><div class="chev-sm">4</div><div class="chev-text"><strong>One next step.</strong></div></div>
]

--

.content-box-yellow[
.font90[`r fontawesome::fa("people-arrows", fill = "#c08a00")` &nbsp; **Swap with a neighboring team** for feedback before you finalize.]
]

.center[`r timer(35L)`]

---
class: segue-yellow

# Scientific Storytelling: <br> Discussion Slides Draft `r top_icon("pen")`

---
### Draft Your Discussion &amp; Conclusion Slides `r fontawesome::fa("clock", fill="#fff", height="0.8em")` 20 min

.pull-left[.font90[
Using the Session 7 **"Presenting Research: The Ending"** Mission Sheet as your guide, draft your **Discussion** and **Conclusion** slides.

Answer three key questions:

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text">What do our results <strong>mean</strong>?</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">Did they <strong>support or challenge</strong> our hypothesis?</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">Who should <strong>care</strong> about this?</div></div>
]]

.pull-right[
.content-box-green[
.font85[`r fontawesome::fa("circle-check", fill = "#009E73")` &nbsp; A strong Discussion slide **adds interpretation** &mdash; it never just repeats what the chart already shows.]
]

.center[`r timer(20L)`]
]

---
class: segue-green

# Homework &amp; After-Class Work `r top_icon("laptop-code")`

---
### After-Class Team Session `r fontawesome::fa("gear", fill="#0148A4", height="0.8em")` 3:15 &ndash; 5:15 PM

.content-box-yellow[
Teams work with a **TA / peer mentor** available. Produce a **tangible output before dinner** &mdash; *not free time.*
]

.font85[
<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>Session 10 Mission Sheet:</strong> write your synthesis &mdash; conclusion sentence, comparison to one literature source, one limitation, one next step.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text"><strong>Draft the Discussion slide:</strong> "What do our results mean?" &mdash; connect each finding to at least one source from your literature review.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text"><strong>DataCORE:</strong> run any remaining analyses; export all final charts as PNG with descriptive names (e.g. <code>fig1_income_health_scatter.png</code>).</div></div>
]

--

.content-box-green[
.font88[**`r fontawesome::fa("flag-checkered", fill = "#009E73")` Team output due before dinner:** draft Discussion slide + completed Session 10 Mission Sheet.]
]

---
### Evening Work Period `r fontawesome::fa("book", fill="#0148A4", height="0.8em")` 7:00 &ndash; 9:00 PM

.pull-left[.font85[
**Individual + partner work** during CTD RTA study hours. TA available for drop-in questions.

- **Individual:** draft your **Discussion paragraph** &mdash; did results support your hypothesis? How do findings compare to the literature? **(200&ndash;250 words)**

- **(RcS):** add a **Discussion/Conclusion** section to your written report with **&ge; 2 literature comparisons (with citations)**. Target **300&ndash;400 words.**
]]

.pull-right[.font85[
.content-box-blue[
**`r fontawesome::fa("users", fill = "#4a7fb5")` Team slide check** *(async is fine):*

- Are all charts **labeled**?
- Does the **story flow**?
- Identify **3 specific things** to fix in tomorrow's session.
]
]]

```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)
to_pdf(from = "10-class.html", to = "10-class.pdf")
```
