---
title: "DataCORE"
subtitle: "Advanced Pandas, Publication-Quality Figures & Tables"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | July 8, 2026</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "60%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
import seaborn as sns
sns.set_theme(style="whitegrid")
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.big-stat { font-size: 2.2rem; font-weight: 700; color: #2e8b57; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Advanced Pandas &mdash; GroupBy, Pivot Tables, Merging <span style="color:#2e8b57;">(1:00&ndash;2:00)</span></h4></li>
    <li><h4>Publication-Quality Figures <span style="color:#2e8b57;">(2:00&ndash;2:45)</span></h4></li>
    <li><h4>Scientific Storytelling &mdash; Table Captions <span style="color:#2e8b57;">(2:45&ndash;3:00)</span></h4></li>
</ol>

.font90[*This morning you met your data. This afternoon you **slice it by groups**, build **publication-ready tables and figures**, and learn to **caption** them like a journal article.*]

---
class: segue-green

# 1:00 &ndash; 2:00 &middot; Advanced Pandas
## .white[GroupBy, Pivot Tables &amp; Merging]

---
### `r fontawesome::fa("arrows-turn-right", height = "1em", fill = "#2e8b57")` From one variable to **many**

.pull-left[
.font90[
- So far: one variable at a time (a mean, a histogram).
- Real analysis asks **comparative** questions: *does the pattern differ **by group**?*
- Three workhorse tools do almost all of it:
]
]

.pull-right[
.content-box-blue[
.font82[
`r fontawesome::fa("layer-group", fill = "#4a7fb5")` **groupby** &mdash; summaries within each group<br>
`r fontawesome::fa("table-cells", fill = "#4a7fb5")` **pivot_table** &mdash; publication-style tables<br>
`r fontawesome::fa("code-merge", fill = "#4a7fb5")` **merge** &mdash; combine two datasets on a key
]
]
]

.font80[**Our data:** the same wage survey &mdash; `education`, `gender`, `region`, `experience_years`, `hourly_wage`.]

---
### `r fontawesome::fa("file-csv", height = "1em", fill = "#2e8b57")` Load the data

.panelset[
.panel[.panel-name[Python]
```{python}
wages = pd.read_csv("wages.csv")
print(wages.head())
print("N =", len(wages))
```
]
.panel[.panel-name[R]
```{r}
wages <- read.csv("wages.csv")
head(wages)
nrow(wages)                                  # pandas: len(wages)
```
]
]

.font80[`r fontawesome::fa("table-columns", fill = "#4a7fb5")` &nbsp; One row per worker. We'll summarise **across** workers, grouped by their characteristics.]

---
### `r fontawesome::fa("layer-group", height = "1em", fill = "#2e8b57")` GroupBy &mdash; multi-level summaries

.panelset[
.panel[.panel-name[Python]
```{python}
print(wages.groupby(["education", "gender"])["hourly_wage"]   # group by TWO keys  #<<
           .agg(["mean", "std", "count"]).round(2))
```
]
.panel[.panel-name[R]
```{r}
wages |>
  group_by(education, gender) |>                              # group by TWO keys  #<<
  summarise(mean = round(mean(hourly_wage), 2),
            std  = round(sd(hourly_wage), 2),
            count = n(), .groups = "drop")
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; Grouping by **two** keys gives one row per *combination*. `agg([...])` computes several statistics at once &mdash; mean, spread, and sample size together.]

---
### `r fontawesome::fa("table-cells", height = "1em", fill = "#2e8b57")` Pivot table &mdash; a publication-style layout

.panelset[
.panel[.panel-name[Python]
```{python}
print(pd.pivot_table(wages, values="hourly_wage",
                     index="education", columns="gender",
                     aggfunc="mean").round(2))
```
]
.panel[.panel-name[R]
```{r}
wages |>
  group_by(education, gender) |>
  summarise(mean_wage = round(mean(hourly_wage), 2), .groups = "drop") |>
  tidyr::pivot_wider(names_from = gender, values_from = mean_wage)
```
]
]

.font80[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; A pivot turns **groups into rows and columns** &mdash; the exact shape of a results table you'd drop into a paper or slide.]

---
### `r fontawesome::fa("code-merge", height = "1em", fill = "#2e8b57")` Merge &mdash; combine two datasets on a key

.panelset[
.panel[.panel-name[Python]
```{python}
region_info = pd.read_csv("region_info.csv")          # cost of living by region
merged = pd.merge(wages, region_info, on="region")    # join on the shared key  #<<
print(merged.head())
```
]
.panel[.panel-name[R]
```{r}
region_info <- read.csv("region_info.csv")            # cost of living by region
merged <- merge(wages, region_info, by = "region")    # join on the shared key  #<<
head(merged)
```
]
]

.font80[`r fontawesome::fa("key", fill = "#4a7fb5")` &nbsp; Merging matches rows by a **shared key** (here `region`) &mdash; the same way you'd join county health data to county income data on a **FIPS code**.]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` Team Task
## .white[Ask Your Data a Grouped Question]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Which groupby matters for **your** question? &mdash; ~12 min

.pull-left[
.font90[
1. Pick the **grouping variable** most relevant to your research question.
2. Run `df.groupby('group')['outcome'].agg(['mean','std','count'])` (Python) **or** the `group_by() |> summarise()` (R).
3. **Stretch:** group by **two** variables, or build a `pivot_table`.
]
]

.pull-right[
.content-box-blue[
.font85[
**Then answer in one sentence:**

*"Grouped by **[variable]**, the **[outcome]** was highest for **[group]** (mean = \_\_) and lowest for **[group]** (mean = \_\_)."*

What does that tell you about your question?
]
]

.center[`r timer(12L)`]
]

---
class: segue-green

# 2:00 &ndash; 2:45 &middot; Publication-Quality Figures
## .white[Charts You'd Show at a Conference]

---
### `r fontawesome::fa("award", height = "1em", fill = "#2e8b57")` One rule

.content-box-red[
.font100[`r fontawesome::fa("triangle-exclamation", fill = "#fff")` &nbsp; **If you would be embarrassed to show this chart in a conference presentation, do not put it in your slides.**]
]

.three-col[
.col[
.content-box-blue[
.font80[
**Readable**<br>
Clear axis labels with units, a title, legible font sizes.
]
]
]
.col[
.content-box-green[
.font80[
**Honest**<br>
Show uncertainty (error bars); don't truncate axes to exaggerate.
]
]
]
.col[
.content-box-yellow[
.font80[
**Accessible**<br>
Colorblind-safe palettes; don't rely on red/green alone.
]
]
]
]

.font80[`r fontawesome::fa("palette", fill = "#4a7fb5")` &nbsp; We'll upgrade a plain chart step by step: **palette &rarr; error bars &rarr; annotation &rarr; export at 300 DPI.**]

---
### `r fontawesome::fa("palette", height = "1em", fill = "#2e8b57")` Colorblind-safe palettes

.panelset[
.panel[.panel-name[Python (seaborn)]
```{python, fig.width=5.6, fig.height=2.8, out.width="48%", results="hide"}
order = wages.groupby("education")["hourly_wage"].mean().sort_values()
colors = sns.color_palette("colorblind", len(order))      # colorblind-safe  #<<
order.plot(kind="bar", color=colors)
plt.ylabel("Mean hourly wage ($)"); plt.xlabel("")
plt.title("Mean wage by education")
plt.xticks(rotation=20); plt.tight_layout(); plt.show()
```
]
.panel[.panel-name[R (ggplot2)]
```{r, fig.width=5.6, fig.height=2.8, out.width="48%"}
wages |> group_by(education) |>
  summarise(mean_wage = mean(hourly_wage)) |>
  ggplot(aes(reorder(education, mean_wage), mean_wage, fill = education)) +
  geom_col() + scale_fill_viridis_d() +                   # colorblind-safe  #<<
  labs(x = NULL, y = "Mean hourly wage ($)", title = "Mean wage by education") +
  theme_minimal() + theme(legend.position = "none")
```
]
]

.font80[`r fontawesome::fa("eye", fill = "#4a7fb5")` &nbsp; `seaborn`'s **"colorblind"** palette and ggplot's **viridis** are designed to stay distinguishable for color-vision deficiencies.]

---
### `r fontawesome::fa("grip-lines-vertical", height = "1em", fill = "#2e8b57")` Error bars &mdash; show the uncertainty

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5.6, fig.height=2.8, out.width="48%", results="hide"}
g = wages.groupby("education")["hourly_wage"]
means = g.mean(); ses = g.std() / np.sqrt(g.count())      # standard error  #<<
means.plot(kind="bar", yerr=ses, capsize=4, color="#4a7fb5")
plt.ylabel("Mean hourly wage ($)"); plt.xlabel("")
plt.title("Mean wage by education (±1 SE)")
plt.xticks(rotation=20); plt.tight_layout(); plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5.6, fig.height=2.8, out.width="48%"}
wages |> group_by(education) |>
  summarise(mean = mean(hourly_wage),
            se = sd(hourly_wage) / sqrt(n())) |>           # standard error  #<<
  ggplot(aes(education, mean)) +
  geom_col(fill = "#4a7fb5") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  labs(x = NULL, y = "Mean hourly wage ($)") + theme_minimal()
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; A bar without error bars hides how **certain** the average is. **mean &plusmn; 1 SE** is the standard add-on.]

---
### `r fontawesome::fa("location-dot", height = "1em", fill = "#2e8b57")` Annotate the point that matters

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5.6, fig.height=2.5, out.width="42%", results="hide"}
x = wages["experience_years"]; y = wages["hourly_wage"]
plt.scatter(x, y, alpha=0.5, color="#4a7fb5")
i = y.idxmax()                                            # highest earner  #<<
plt.annotate("highest earner", (x[i], y[i]), xytext=(x[i]-13, y[i]+1),
             arrowprops=dict(arrowstyle="->", color="#D55E00"))
plt.xlabel("Years of experience"); plt.ylabel("Hourly wage ($)")
plt.tight_layout(); plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5.6, fig.height=2.5, out.width="42%"}
i <- which.max(wages$hourly_wage)                          # highest earner  #<<
plot(wages$experience_years, wages$hourly_wage, col = "#4a7fb5",
     xlab = "Years of experience", ylab = "Hourly wage ($)")
points(wages$experience_years[i], wages$hourly_wage[i], col = "#D55E00", pch = 19)
text(wages$experience_years[i], wages$hourly_wage[i],
     "highest earner", pos = 2, col = "#D55E00")
```
]
]

.font80[`r fontawesome::fa("comment", fill = "#4a7fb5")` &nbsp; One well-placed label guides the eye to your **key data point** &mdash; far better than making the viewer hunt for it.]

---
### `r fontawesome::fa("floppy-disk", height = "1em", fill = "#2e8b57")` Export at 300 DPI for slides &amp; print

.panelset[
.panel[.panel-name[Python]
```{python, eval=FALSE}
plt.savefig("fig1_wage_by_education.png",
            dpi=300, bbox_inches="tight")     # crisp; no clipped labels  #<<
```
]
.panel[.panel-name[R]
```{r, eval=FALSE}
ggsave("fig1_wage_by_education.png",
       width = 6, height = 4, dpi = 300)       # crisp; sized in inches  #<<
```
]
]

.content-box-yellow[
.font85[`r fontawesome::fa("file-image", fill = "#c08a00")` &nbsp; Give every figure a **descriptive file name** (`fig1_wage_by_education.png`, not `Untitled.png`) and a **caption**. `bbox_inches="tight"` stops axis labels from being cut off.]
]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` Team Task
## .white[Finalize Your Summit Figures]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Your 2&ndash;3 best charts &mdash; ~12 min

.pull-left[
.font90[
1. Choose the **2&ndash;3 charts** that best answer your research question.
2. Upgrade each: **colorblind-safe palette**, clear labels with **units**, a title.
3. Add **error bars** or **annotate** the key point where it helps.
4. **Save** each at **300 DPI** with a descriptive file name.
]
]

.pull-right[
.content-box-blue[
.font85[
**Each chart needs:**

- A descriptive **file name**
- A one-line **caption**
- A reason it's in the deck &mdash; *what question does it answer?*
]
]

.center[`r timer(12L)`]
]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[Table Captions]

---
### `r fontawesome::fa("table-list", height = "1em", fill = "#2e8b57")` A table is useless without a caption

.panelset[
.panel[.panel-name[Python]
```{python}
print(wages.groupby("education")["hourly_wage"]
           .agg(["mean", "median", "std", "min", "max"]).round(2))
```
]
.panel[.panel-name[R]
```{r}
wages |> group_by(education) |>
  summarise(mean = round(mean(hourly_wage), 2),
            median = median(hourly_wage),
            std = round(sd(hourly_wage), 2),
            min = min(hourly_wage), max = max(hourly_wage))
```
]
]

.font80[`r fontawesome::fa("circle-question", fill = "#4a7fb5")` &nbsp; A reader who sees only this table should still know **what it shows, where it came from, and one thing to notice.** That's the caption's job.]

---
### `r fontawesome::fa("feather-pointed", height = "1em", fill = "#2e8b57")` APA table format

.content-box-green[
.font88[`r fontawesome::fa("quote-left", fill = "#2e8b57")` &nbsp; **Table 1.** *Hourly wages by education level (N = 220 workers).* **Note.** Data from the 2023 wage survey; values are in U.S. dollars per hour. Mean hourly wage rises with education, from \$22 (high school) to \$35 (graduate).]
]

.pull-left[
.content-box-blue[
.font80[
**The shape:**<br>
*"Table 1. **[Descriptive title]**. Note. **[Source and any key definitions]**."*
]
]
]

.pull-right[
.content-box-yellow[
.font80[
**Why it matters:**<br>
APA table format is the **standard** in the social and health sciences &mdash; titles above the table, notes below.
]
]
]

---
### `r fontawesome::fa("pen-fancy", height = "1em", fill = "#2e8b57")` Write your table's caption &mdash; now

.pull-left[
.font90[
Write a caption for **your** summary-statistics table:

.content-box-blue[
.font80[*"Table 1. **[descriptive title]**. Note. **[source, sample, definitions]**; **[one notable pattern]**."*]
]

Read it to a partner who hasn't seen your data.
]
]

.pull-right[
.content-box-yellow[
.font85[
`r fontawesome::fa("circle-check", fill = "#c08a00")` **Checklist:**

- Descriptive **title** (not just "Table 1")?
- **Sample / N** stated?
- **Source &amp; year**?
- Units / key **definitions**?
- **One pattern** worth noticing?
]
]

.center[`r timer(10L)`]
]

---
class: segue-green

# .white[You can now group, pivot, merge, and visualize.]
## .white[Publication-quality means readable, honest, and accessible.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)
to_pdf(from = "08-pm-datacore.html", to = "08-pm-datacore.pdf")
```
