---
title: "DataCORE"
subtitle: "Descriptive Statistics, Distributions & Visualizing Your Variables"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "62%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.stat-card {
  background: #eaf3fb;
  border-radius: 8px;
  padding: 0.5em 0.2em;
  text-align: center;
}
.stat-card .big { font-size: 1.6em; font-weight: bold; color: #2e8b57; }
.stat-card .lbl { font-size: 0.8em; color: #333; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Descriptive Statistics &amp; Distributions</h4></li>
    <li><h4>Visualizing Your Variables &mdash; Histogram, Bar, Scatter</h4></li>
    <li><h4>Scientific Storytelling &mdash; Chart Annotations</h4></li>
</ol>

.font90[*Yesterday you computed one number for your DV. Today we ask: what is the **shape** of your data &mdash; and how do we **show** it?*]

---
class: segue-green

# 1:00 &ndash; 1:50 &middot; Descriptive Statistics & Distributions
## .white[From a Single Number to a Distribution]

---
### `r fontawesome::fa("ruler", height = "1em", fill = "#2e8b57")` One number hides the story

.pull-left[
.font90[
- A **mean** tells you the *center* &mdash; but two very different datasets can share the same mean.

- We need numbers for the **center** *and* the **spread**, plus a picture of the **shape**.

- Today's pipeline: **summarise &rarr; visualize &rarr; describe in plain English.**
]
]

.pull-right[
.content-box-blue[
.font90[**Big idea:** a *distribution* is the full pattern of values a variable takes &mdash; not just its average.]
]
]

.font80[`r fontawesome::fa("table", fill = "#4a7fb5")` &nbsp; We use `study_habits.csv` &mdash; swap in **your** team file.]

---
### `r fontawesome::fa("file-csv", height = "1em", fill = "#2e8b57")` Reload the dataset

.panelset[
.panel[.panel-name[Python]
```{python}
import pandas as pd
df = pd.read_csv("study_habits.csv")   # rows = students, columns = variables   #<<
print(df.head())
```
]
.panel[.panel-name[R]
```{r}
library(tidyverse)
df <- read_csv("study_habits.csv")     # rows = students, columns = variables   #<<
head(df)
```
]
]

.font80[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; Columns: `grade_level`, `study_method`, `hours_of_sleep`, `test_score`.]

---
### `r fontawesome::fa("calculator", height = "1em", fill = "#2e8b57")` Mean, median, standard deviation

.panelset[
.panel[.panel-name[Python]
```{python}
dv = df["test_score"]
print("mean  :", round(dv.mean(), 2))     # the average                          #<<
print("median:", dv.median())             # the middle value                     #<<
print("std   :", round(dv.std(), 2))      # typical distance from the mean        #<<
```
]
.panel[.panel-name[R]
```{r}
dv <- df$test_score
cat("mean  :", round(mean(dv, na.rm = TRUE), 2), "\n")   # the average            #<<
cat("median:", median(dv, na.rm = TRUE), "\n")           # the middle value       #<<
cat("std   :", round(sd(dv, na.rm = TRUE), 2), "\n")     # spread                 #<<
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; `df.describe()` (pandas) / `summary(df)` (R) gives all of these at once.]

---
### `r fontawesome::fa("list-check", height = "1em", fill = "#2e8b57")` Categorical stats: frequencies and percentages

.panelset[
.panel[.panel-name[Python]
```{python}
cat_counts = df["study_method"].value_counts(dropna=False).rename("frequency")
cat_percent = (df["study_method"].value_counts(dropna=False, normalize=True) * 100).round(1).rename("percent")
cat_table = pd.concat([cat_counts, cat_percent], axis=1)
print(cat_table)
```
]
.panel[.panel-name[R]
```{r}
cat_table <- df %>%
  count(study_method, name = "frequency") %>%
  mutate(percent = round(100 * frequency / sum(frequency), 1))

cat_table
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; Use this for any categorical variable (e.g., `grade_level`, `study_method`) to report how many observations are in each group and what share of the sample each group represents.]

---
### `r fontawesome::fa("scale-unbalanced", height = "1em", fill = "#2e8b57")` What each number tells you

.three-col[
.col[
.content-box-blue[
.font80[
**Mean**<br>
The balance point. Uses *every* value &mdash; so it gets **pulled by outliers**.
]
]
]
.col[
.content-box-green[
.font80[
**Median**<br>
The middle value. **Robust to outliers** &mdash; better when data is skewed.
]
]
]
.col[
.content-box-yellow[
.font80[
**Std. Dev.**<br>
Typical distance from the mean. **Small** = clustered; **large** = spread out.
]
]
]
]

.content-box-blue[
.font85[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; **When is median better than mean?** Income, house prices, test scores with a few extreme values &mdash; one billionaire (or one zero) drags the mean, not the median.]
]

---
### `r fontawesome::fa("chart-simple", height = "1em", fill = "#2e8b57")` `df.hist()` &mdash; the shape of your distribution

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5, fig.height=2.5, out.width="45%", results="hide"}
df["test_score"].hist(bins=8, color="#2e8b57", edgecolor="white")
plt.title("Distribution of Test Scores")
plt.xlabel("Test score"); plt.ylabel("Number of students")
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.5, out.width="45%"}
ggplot(df, aes(test_score)) +
  geom_histogram(bins = 8, fill = "#2e8b57", color = "white") +
  labs(title = "Distribution of Test Scores",
       x = "Test score", y = "Number of students")
```
]
]

.font80[`r fontawesome::fa("eye", fill = "#4a7fb5")` &nbsp; **Read the shape:** is it *normal* (bell), *skewed* (a long tail), or *bimodal* (two humps)?]

---
### `r fontawesome::fa("shapes", height = "1em", fill = "#2e8b57")` Three shapes you will meet

.three-col[
.col[
.content-box-blue[
.font80[
**Normal**<br>
Symmetric bell. Mean &asymp; median. Most values near the center.
]
]
]
.col[
.content-box-yellow[
.font80[
**Skewed**<br>
One long tail. Mean is pulled toward the tail &mdash; **report the median**.
]
]
]
.col[
.content-box-green[
.font80[
**Bimodal**<br>
Two peaks. Often a sign of **two hidden groups** mixed together.
]
]
]
]

.font80[`r fontawesome::fa("users", fill = "#4a7fb5")` &nbsp; A bimodal DV is a clue: maybe boys vs. girls, or tutored vs. solo studiers behave differently.]

---
### `r fontawesome::fa("box", height = "1em", fill = "#2e8b57")` `df.boxplot()` &mdash; spot outliers visually

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5, fig.height=2.5, out.width="45%", results="hide"}
df.boxplot(column="test_score")
plt.title("Test Scores -- Boxplot")
plt.ylabel("Test score")
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.5, out.width="45%"}
ggplot(df, aes(y = test_score)) +
  geom_boxplot(fill = "#2e8b57", alpha = 0.6) +
  labs(title = "Test Scores -- Boxplot", y = "Test score")
```
]
]

.font80[`r fontawesome::fa("circle-dot", fill = "#4a7fb5")` &nbsp; The **box** holds the middle 50%; the **line** is the median; **dots** beyond the whiskers are possible outliers.]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` Team Pipeline
## .white[Describe Your DV's Distribution]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Run the pipeline on your DV &mdash; ~15 min

.pull-left[
.font90[
**On your team's dependent variable, record:**

<div class="three-col" style="margin-top:0.6em;">
  <div class="stat-card"><div class="lbl">min</div><div class="big">?</div></div>
  <div class="stat-card"><div class="lbl">max</div><div class="big">?</div></div>
  <div class="stat-card"><div class="lbl">mean</div><div class="big">?</div></div>
  <div class="stat-card"><div class="lbl">SD</div><div class="big">?</div></div>
</div>

Then run `df.hist()` **and** `df.boxplot()` on it.
]
]

.pull-right[
.content-box-yellow[
.font85[
**Write ONE sentence** describing the distribution in plain English:

*"Test scores ranged from 65 to 96, centered around 81, and were roughly **bell-shaped** with no extreme outliers."*
]
]
]

.font80[`r fontawesome::fa("pen", fill = "#4a7fb5")` &nbsp; Record the four numbers **and** your sentence on your Mission Sheet.]

---
class: segue-green

# 1:50 &ndash; 2:45 &middot; Visualizing Your Variables
## .white[The Right Chart for the Right Variable]

---
### `r fontawesome::fa("chart-line", height = "1em", fill = "#2e8b57")` Match the chart to the variable

.three-col[
.col[
.content-box-blue[
.font85[
.center[`r fontawesome::fa("chart-simple", height = "1.6em", fill = "#0148A4")`]

**Histogram**<br>
One *continuous* variable. **Shows its distribution.**
]
]
]
.col[
.content-box-green[
.font85[
.center[`r fontawesome::fa("chart-column", height = "1.6em", fill = "#2e8b57")`]

**Bar chart**<br>
One *categorical* variable. **Compares groups.**
]
]
]
.col[
.content-box-yellow[
.font85[
.center[`r fontawesome::fa("braille", height = "1.6em", fill = "#c08a00")`]

**Scatter plot**<br>
*Two continuous* variables. **Shows their relationship.**
]
]
]
]

.font80[`r fontawesome::fa("hand-point-right", fill = "#4a7fb5")` &nbsp; Instructor demos one; you replicate all three for **your** variables.]

---
### `r fontawesome::fa("chart-simple", height = "1em", fill = "#2e8b57")` Histogram &mdash; a continuous variable

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5, fig.height=2.5, out.width="45%", results="hide"}
df["hours_of_sleep"].hist(bins=7, color="#4a7fb5", edgecolor="white")
plt.title("Distribution of Hours of Sleep")
plt.xlabel("Hours of sleep"); plt.ylabel("Number of students")
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.5, out.width="45%"}
ggplot(df, aes(hours_of_sleep)) +
  geom_histogram(bins = 7, fill = "#4a7fb5", color = "white") +
  labs(title = "Distribution of Hours of Sleep",
       x = "Hours of sleep", y = "Number of students")
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; **Bins** group the values &mdash; too few hides shape, too many makes it noisy.]

---
### `r fontawesome::fa("chart-column", height = "1em", fill = "#2e8b57")` Bar chart &mdash; a categorical variable

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5, fig.height=2.5, out.width="45%", results="hide"}
counts = df["study_method"].value_counts()
counts.plot(kind="bar", color="#2e8b57", edgecolor="white")
plt.title("Students by Study Method")
plt.xlabel("Study method"); plt.ylabel("Number of students")
_ = plt.xticks(rotation=0)
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.5, out.width="45%"}
ggplot(df, aes(study_method)) +
  geom_bar(fill = "#2e8b57", color = "white") +
  labs(title = "Students by Study Method",
       x = "Study method", y = "Number of students")
```
]
]

.font80[`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` &nbsp; A **bar chart** counts *categories*; a **histogram** bins *numbers*. Don't mix them up.]

---
### `r fontawesome::fa("braille", height = "1em", fill = "#2e8b57")` Scatter plot &mdash; two continuous variables (IV vs DV)

.panelset[
.panel[.panel-name[Python]
```{python, fig.width=5, fig.height=2.5, out.width="45%", results="hide"}
_ = plt.scatter(df["hours_of_sleep"], df["test_score"], color="#2e8b57")
plt.title("Sleep vs. Test Score")
plt.xlabel("Hours of sleep (IV)"); plt.ylabel("Test score (DV)")
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.5, out.width="45%"}
ggplot(df, aes(hours_of_sleep, test_score)) +
  geom_point(color = "#2e8b57", size = 2) +
  labs(title = "Sleep vs. Test Score",
       x = "Hours of sleep (IV)", y = "Test score (DV)")
```
]
]

.font80[`r fontawesome::fa("arrow-trend-up", fill = "#4a7fb5")` &nbsp; Each dot is one student. Look for a **trend**: do points drift up, down, or scatter randomly?]

---
### `r fontawesome::fa("triangle-exclamation", height = "1em", fill = "#2e8b57")` The one rule for every chart

.pull-left[
.content-box-red[
.font100[
**A chart without a title and axis labels is not a chart &mdash; it is a picture.**
]
]

.font85[Every chart must be **self-explanatory**: a reader should understand it without you standing next to it.]
]

.pull-right[
.content-box-green[
.font85[
**Every chart needs:**

- `r fontawesome::fa("check", fill="#2e8b57")` A clear **title**
- `r fontawesome::fa("check", fill="#2e8b57")` A labeled **x-axis** (with units)
- `r fontawesome::fa("check", fill="#2e8b57")` A labeled **y-axis** (with units)
- `r fontawesome::fa("check", fill="#2e8b57")` Readable text & legend (if needed)
]
]
]

.font80[`r fontawesome::fa("scissors", fill = "#4a7fb5")` &nbsp; If you screenshot a chart with no labels for your portfolio, you've saved a picture &mdash; not a finding.]

---
### `r fontawesome::fa("robot", height = "1em", fill = "#2e8b57")` AI prompt &mdash; generate, run, then polish

.pull-left[
.font90[
**1.** Ask ChatGPT:

.content-box-blue[.font80[*"Write Python code for a **scatter plot** of **[IV]** vs **[DV]** from my pandas DataFrame `df`."*]]

**2.** **Run** it on your real data.

**3.** **Add** color, a title, and axis labels.

**4.** **Screenshot** it for your portfolio.
]
]

.pull-right[
.content-box-yellow[
.font85[
**AI almost always forgets:**

- A descriptive **title**
- **Axis labels** with units
- Sensible **color** & point size
- Handling **missing values**

*You* are the one who makes it self-explanatory.
]
]
]

.font80[`r fontawesome::fa("wand-magic-sparkles", fill = "#4a7fb5")` &nbsp; AI draws fast &mdash; **you** turn its picture into a chart.]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` Team Challenge
## .white[Build All Three Charts on Your Data]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Build &amp; label &mdash; ~20 min

.pull-left[
.font90[
**Using your team data, build:**

<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">1.</strong> A <strong>histogram</strong> of one continuous variable.
</div>
<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">2.</strong> A <strong>bar chart</strong> of one categorical variable.
</div>
<div class="chev-row-sm" style="display:flex;align-items:center;gap:0.6em;margin-bottom:0.4em;">
  <strong style="color:#2e8b57;">3.</strong> A <strong>scatter plot</strong> of your <strong>IV vs DV</strong>.
</div>
]
]

.pull-right[
.content-box-blue[
.font85[
**Every chart must pass the checklist:**

- Title `r fontawesome::fa("check", fill="#2e8b57")`
- X-axis label `r fontawesome::fa("check", fill="#2e8b57")`
- Y-axis label `r fontawesome::fa("check", fill="#2e8b57")`

Screenshot each one for your **portfolio**.
]
]
]

.font80[`r fontawesome::fa("people-group", fill = "#4a7fb5")` &nbsp; Instructor checks that each chart is **self-explanatory** before you save it.]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[Chart Annotations]

---
### `r fontawesome::fa("comment-dots", height = "1em", fill = "#2e8b57")` One sentence under every chart

.pull-left[
.font90[
Add a **one-sentence annotation** below each chart that answers:

- What **pattern** does this show?
- What would a **reader** need to know to interpret it?
]

.content-box-blue[
.font85[`r fontawesome::fa("seedling", fill = "#4a7fb5")` &nbsp; These annotations become the **captions in your Results section.**]
]
]

.pull-right[
.content-box-green[
.font85[
`r fontawesome::fa("eye", fill="#2e8b57")` **Describe** (Results)<br>
*"The bars show higher values in Group A."*
]
]

.content-box-yellow[
.font85[
`r fontawesome::fa("brain", fill="#c08a00")` **Interpret** (Discussion)<br>
*"Group A reports higher scores, which may reflect..."*
]
]
]

---
### `r fontawesome::fa("scale-balanced", height = "1em", fill = "#2e8b57")` Describe now &mdash; interpret later

.pull-left[
.content-box-green[
.font85[
**Describing = Results**

- States *what the chart shows*
- Just the facts: direction, size, groups
- No "why," no causes
- *"Scores rise as sleep increases."*
]
]
]

.pull-right[
.content-box-yellow[
.font85[
**Interpreting = Discussion**

- Explains *what it might mean*
- Offers reasons, caveats, theory
- Saved for **later** in the paper
- *"...perhaps because rested students focus better."*
]
]
]

.content-box-blue[
.font85[`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` &nbsp; **Results are descriptive.** Save the "why" for the Discussion &mdash; mixing them is the most common beginner mistake.]
]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Annotate your three charts &mdash; quick round

.font90[
For **each** of your three charts, write a one-sentence **descriptive** annotation. Then read your scatter-plot annotation aloud to a partner.
]

.content-box-blue[
.font85[
**Template:** *"This [chart type] of [variable(s)] shows [pattern], which means a reader can see [what it tells them]."*
]
]

.content-box-yellow[
.font85[`r fontawesome::fa("circle-check", fill = "#4a7fb5")` &nbsp; Did you stay **descriptive**? If you wrote the word *"because,"* you slipped into interpretation &mdash; move it to the Discussion.]
]

---
class: segue-green

# .white[From numbers to pictures to a story.]
## .white[Tomorrow: comparing groups & testing relationships.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-04-literature-review/pm/04-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-04-literature-review/pm/04-pm-datacore.pdf")
to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-04-literature-review/pm/04-pm-datacore.html")
```
