---
title: "DataCORE"
subtitle: "Parallel Tracks — Core Analysis & Multiple Regression"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | July 9, 2026</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "60%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
import seaborn as sns
from scipy import stats
import statsmodels.formula.api as smf
import statsmodels.api as sm
sns.set_theme(style="whitegrid")
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.big-stat { font-size: 2.2rem; font-weight: 700; color: #2e8b57; }
.chev-row { display: flex; align-items: center; gap: 0.9em; margin-bottom: 0.6em; }
.chev { flex: 0 0 60px; height: 74px; background: #cfe8d8;
  clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
  display: flex; align-items: center; justify-content: center;
  font-weight: bold; font-size: 1.6em; color: #1a4731; }
.chev-row-sm { display: flex; align-items: center; gap: 0.8em; margin-bottom: 0.3em; }
.chev-sm { flex: 0 0 50px; height: 56px; background: #cfe8d8;
  clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
  display: flex; align-items: center; justify-content: center;
  font-weight: bold; color: #1a4731; }
.chev-text { flex: 1; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Multiple Regression + Interpretation <span style="color:#2e8b57;">(1:00&ndash;2:45)</span></h4></li>
    <li><h4>Scientific Storytelling &mdash; Figure + Interpretation Pairing <span style="color:#2e8b57;">(2:45&ndash;3:00)</span></h4></li>
</ol>

.font90[*Two tracks run **in parallel** today. **Foundation** builds the core chart-and-statistic for your question; **Scholars** fits a multiple regression and interprets it carefully. Both share the same data and the same closing exercise.*]

---
### `r fontawesome::fa("code-branch", height = "1em", fill = "#2e8b57")` Parallel tracks &mdash; pick your lane

.pull-left[
.content-box-blue[
**`r fontawesome::fa("seedling", fill = "#4a7fb5")` Foundation (RcF)**

One clear chart + one key statistic, interpreted in plain language. The goal: a **defensible finding** you fully understand.
]
]

.pull-right[
.content-box-green[
**`r fontawesome::fa("graduation-cap", fill = "#009E73")` Scholars (RcS)**

Multiple regression with 2+ predictors, full output table, assumption checks, and a **"holding constant"** interpretation.
]
]

<div style="clear: both;"></div>

.font80[**Shared dataset:** the wage survey &mdash; `education`, `gender`, `region`, `experience_years`, `hourly_wage`.]

```{python load-data, include=FALSE}
wages = pd.read_csv("wages.csv")
```

```{r load-data-r, include=FALSE}
wages <- read.csv("wages.csv")
```

---
class: segue-green

# 1:00 &ndash; 2:45 &middot; Scholars Track
## .white[Multiple Regression &amp; Interpretation]

---
### `r fontawesome::fa("layer-group", height = "1em", fill = "#2e8b57")` Why **multiple** regression?

.pull-left[
.font90[
- Simple regression: one predictor (wage ~ experience).
- But wage also depends on **education**. If experienced workers are also more educated, simple regression **confounds** the two.
- Multiple regression estimates each predictor's effect **holding the others constant.**
]
]

.pull-right[
.content-box-blue[
.font85[`r fontawesome::fa("equals", fill = "#4a7fb5")` &nbsp; **wage = &beta;&#8320; + &beta;&#8321;&middot;experience + &beta;&#8322;&middot;education**

Each **&beta;** is a *partial* slope &mdash; the effect of that variable with the rest held fixed.]
]
]

.font80[**DV** = hourly wage &middot; **IVs** = years of experience **and** education level.]

---
### `r fontawesome::fa("table", height = "1em", fill = "#2e8b57")` Fit OLS with 2+ predictors

.panelset[
.panel[.panel-name[Python (statsmodels)]
```{python}
model = smf.ols("hourly_wage ~ experience_years + C(education)", data=wages).fit()
print(model.summary().tables[1])
```
]
.panel[.panel-name[R (base)]
```{r}
model <- lm(hourly_wage ~ experience_years + education, data = wages)
round(summary(model)$coefficients, 4)
```
]
]

.font80[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; `C(education)` / a factor adds one row **per category** (vs. a baseline). Each coefficient is that group's gap from the baseline, **holding experience constant.**]

---
### `r fontawesome::fa("list-ol", height = "1em", fill = "#2e8b57")` Read the full output

.panelset[
.panel[.panel-name[Python]
```{python}
print(f"R-squared     = {model.rsquared:.3f}")
print(f"Adj. R-squared= {model.rsquared_adj:.3f}")
print(f"F-statistic   = {model.fvalue:.1f}   (p = {model.f_pvalue:.2e})")
print(f"N             = {int(model.nobs)}")
```
]
.panel[.panel-name[R]
```{r}
s <- summary(model)
cat(sprintf("R-squared      = %.3f\n", s$r.squared))
cat(sprintf("Adj. R-squared = %.3f\n", s$adj.r.squared))
cat(sprintf("F-statistic    = %.1f (p = %.2e)\n",
            s$fstatistic[1],
            pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE)))
```
]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; **Each coefficient:** the effect of that variable. **Std err / t / p:** is it distinguishable from 0? **R&sup2;:** variation explained. **F-stat:** is the model as a whole useful?]

---
### `r fontawesome::fa("comment", height = "1em", fill = "#2e8b57")` Say it the careful way

```{python extract-coef, include=FALSE}
b_exp = float(model.params["experience_years"])
p_exp = float(model.pvalues["experience_years"])
```

.content-box-green[
.font88[`r fontawesome::fa("quote-left", fill = "#2e8b57")` &nbsp; *"**Holding education constant,** each additional year of experience is associated with a **\$`r sprintf("%.2f", py$b_exp)`** change in hourly wage (p `r if (py$p_exp < 0.001) "< 0.001" else paste("=", sprintf("%.3f", py$p_exp))`)."*]
]

.pull-left[
.content-box-blue[
.font80[
**The template:**<br>
*"Holding **[controls]** constant, a one-unit increase in **[IV]** is associated with a **[coef]** unit change in **[DV]** (p = \_\_)."*
]
]
]

.pull-right[
.content-box-yellow[
.font80[
**"Associated with," not "causes."**<br>
Even with controls, this is observational &mdash; unmeasured factors may remain.
]
]
]

---
### `r fontawesome::fa("ruler-combined", height = "1em", fill = "#2e8b57")` Check the assumptions

.panelset[
.panel[.panel-name[Residuals · Python]
```{python, fig.width=5.4, fig.height=2.6, out.width="44%", results="hide"}
plt.scatter(model.fittedvalues, model.resid, alpha=0.5, color="#4a7fb5")
plt.axhline(0, color="#D55E00", lw=2)
plt.xlabel("Fitted values"); plt.ylabel("Residuals")
plt.title("Residuals vs. fitted")
plt.tight_layout(); plt.show()
```
]
.panel[.panel-name[Q-Q · Python]
```{python, fig.width=5.4, fig.height=2.6, out.width="44%", results="hide"}
sm.qqplot(model.resid, line="s")
plt.title("Normal Q-Q plot")
plt.tight_layout(); plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=6.4, fig.height=2.7, out.width="62%"}
par(mfrow = c(1, 2))
plot(model, which = 1)   # residuals vs fitted
plot(model, which = 2)   # normal Q-Q
```
]
]

.font80[`r fontawesome::fa("eye", fill = "#4a7fb5")` &nbsp; **Residuals vs. fitted:** want a shapeless cloud around 0 (no funnel, no curve). **Q-Q plot:** points near the line = roughly normal residuals.]

---
### `r fontawesome::fa("clipboard-check", height = "1em", fill = "#2e8b57")` Scholars Mission Sheet

.content-box-green[
Complete the **Session 9 Scholars Mission Sheet** on your own data.
]

.font88[
<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text">Choose a <strong>variable pair</strong> (DV + key IV) plus at least one <strong>control</strong>.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">State the <strong>relationship found</strong> &mdash; coefficient, sign, significance.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">Write the <strong>plain-language</strong> "holding constant" interpretation.</div></div>
<div class="chev-row-sm"><div class="chev-sm">4</div><div class="chev-text">Name <strong>one causal limitation</strong> &mdash; what could still confound it?</div></div>
]

.center[`r timer(45L)`]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[Figure + Interpretation Pairing]

---
### `r fontawesome::fa("link", height = "1em", fill = "#2e8b57")` Make them a matched pair

.pull-left[
.font90[
A great results slide is a **chart + one sentence** that fit together perfectly.

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text">Write a <strong>1-sentence interpretation</strong> of your strongest chart.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text">Read it aloud <strong>without</strong> showing the chart &mdash; does it stand on its own?</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">Now show the chart &mdash; does it <strong>prove</strong> the sentence?</div></div>
]
]

.pull-right[
.content-box-blue[
.font82[
**The interpretation tells the audience what to see; the chart proves it.**

*Example:* the sentence *"Wages rise steadily with experience"* paired with a scatter that clearly trends up.
]
]

.center[`r timer(15L)`]
]

---
class: segue-green

# .white[Foundation or Scholars, the discipline is the same:]
## .white[a clear finding, an honest interpretation, a chart that proves it.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)

to_pdf(from = "09-pm-datacore.html",
       to = "09-pm-datacore.pdf")
```
