---
title: "DataCORE"
subtitle: "Exploring Data with Code"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE)
```

```{python py-setup, include=FALSE}
import pandas as pd
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Orientation &mdash; Tools & Environment</h4></li>
    <li><h4>Exploring Your First Dataset &mdash; NBA Home Advantage</h4></li>
    <li><h4>AI as a Coding Assistant</h4></li>
    <li><h4>Scientific Storytelling &mdash; Data Captions</h4></li>
</ol>

.font90[*DataCORE isn't about software &mdash; it's about asking questions and finding answers in data.*]

---
class: segue-green

# 1:00 &ndash; 1:20 &middot; Orientation
## Tools & Environment

---
### `r fontawesome::fa("laptop-code", height = "1em", fill = "#2e8b57")` DataCORE Orientation

.pull-left[
.font90[
- Everything runs in **Google Colab**, in your browser &mdash; **no installation**.

- You'll **open a notebook**, **run a cell**, and see output instantly.

- Today's first cell just says hello `r fontawesome::fa("arrow-right", fill = "#4a7fb5")`
]

.content-box-blue[
.font90[**Big idea:** DataCORE is *not* about software. It's about **asking questions and finding answers in data**.]
]
]

.pull-right[
.panelset[
.panel[.panel-name[Python]
```{python}
print("Hello, Research!")   #<<
```
]
.panel[.panel-name[R]
```{r}
print("Hello, Research!")   #<<
```
]
]
]

---
class: segue-green

# 1:20 &ndash; 2:20 &middot; Exploring Your First Dataset
## NBA Home Advantage

---
### `r fontawesome::fa("table", height = "1em", fill = "#2e8b57")` Step 1 &middot; Load & peek at the data

.font90[Open the shared notebook, then **look before you leap** &mdash; preview the rows and the shape.]

.panelset[
.panel[.panel-name[Python]
```{python}
import pandas as pd
df = pd.read_csv("nba_home_away.csv")
print(df.head())       # first 5 rows   #<<
```

```{python}
df.shape        # rows = games, cols = fields   #<<
```
]
.panel[.panel-name[R]
```{r}
library(tidyverse)
df <- read_csv("nba_home_away.csv")
head(df)        # first 5 rows   #<<
```

```{r}
dim(df)         # rows = games, cols = fields   #<<
```
]
]

.font80[`r fontawesome::fa("circle-question", fill = "#4a7fb5")` &nbsp; How many games are in the data? How many teams?]

---
### `r fontawesome::fa("calculator", height = "1em", fill = "#2e8b57")` Step 2 &middot; Basic statistics at a glance

.panelset[
.panel[.panel-name[Python]
```{python}
print(df.describe().round(2))    # summary stats per column   #<<
```
]
.panel[.panel-name[R]
```{r}
summary(df)      # min, median, mean, max for every column   #<<
```
]
]

.font80[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; One line gives you the **shape of the whole dataset** &mdash; means, spread, and extremes.]

---
### `r fontawesome::fa("percent", height = "1em", fill = "#2e8b57")` Step 3 &middot; Calculate the home win rate

.font90[`home_wins` is **1** when the home team won and **0** when it lost &mdash; so its **average is the win rate**.]

.panelset[
.panel[.panel-name[Python]
```{python}
df["home_wins"].mean()    # share of games won by the home team   #<<
```
]
.panel[.panel-name[R]
```{r}
mean(df$home_wins)        # share of games won by the home team   #<<
```
]
]

.font80[`r fontawesome::fa("basketball", fill = "#4a7fb5")` &nbsp; Is it above 0.50? By how much? That's the **home advantage** in one number.]

---
### `r fontawesome::fa("chart-column", height = "1em", fill = "#2e8b57")` Step 4 &middot; Make a bar chart

.panelset[
.panel[.panel-name[Python]
```{python, out.width="58%"}
import matplotlib.pyplot as plt

home = df["home_wins"].mean()
away = 1 - home

plt.figure(figsize=(6, 2.4))
plt.bar(["Home", "Away"], [home, away],
        color=["#0148A4", "#9bc3e6"])    # color the bars   #<<
plt.ylabel("Win rate")
plt.title("NBA: Home vs. Away Win Rate")  # title & labels   #<<
plt.show()
```
]
.panel[.panel-name[R]
```{r, fig.width=5, fig.height=2.0, out.width="44%"}
rates <- tibble(
  side = c("Home", "Away"),
  rate = c(mean(df$home_wins), 1 - mean(df$home_wins))
)

ggplot(rates, aes(side, rate, fill = side)) +
  geom_col() +                            # the bars       #<<
  scale_fill_manual(values = c("#9bc3e6", "#0148A4")) +
  labs(y = "Win rate",
       title = "NBA: Home vs. Away Win Rate")  # title & labels   #<<
```
]
]

.font80[`r fontawesome::fa("pen", fill = "#4a7fb5")` &nbsp; On your **Mission Sheet:** *What surprised me in the data?*]

---
class: segue-green

# 2:20 &ndash; 2:45 &middot; AI as a Coding Assistant

---
### `r fontawesome::fa("robot", height = "1em", fill = "#2e8b57")` You asked the AI for a "win percentage" column...

.font90[It instantly wrote this. You run it &mdash; and the numbers look **way too high** (over 100%!). Can you spot the bug?]

.panelset[
.panel[.panel-name[Python]
```{python}
# AI: "what % of games did the home team win?"
home_win_pct = df["home_wins"].sum() / df["away_wins"].sum() * 100   # ÷ away wins   #<<
home_win_pct
```
]
.panel[.panel-name[R]
```{r}
# AI: "what % of games did the home team win?"
home_win_pct <- sum(df$home_wins) / sum(df$away_wins) * 100   # / away wins   #<<
home_win_pct
```
]
]

.font80[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; That gives **139.8%** &mdash; home wins should be divided by **total games**, not away wins.]

---
### `r fontawesome::fa("circle-check", height = "1em", fill = "#2e8b57")` Fix it: divide by total games

.panelset[
.panel[.panel-name[Python]
```{python}
home_win_pct = df["home_wins"].sum() / len(df) * 100   #<<
home_win_pct
```
]
.panel[.panel-name[R]
```{r}
home_win_pct <- sum(df$home_wins) / nrow(df) * 100   #<<
home_win_pct
```
]
]

.content-box-blue[
.font110[**Key lesson:** AI *writes* code &mdash; **humans verify it**. You have to understand what the code does.]
]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## Data Captions

---
### `r fontawesome::fa("comment-dots", height = "1em", fill = "#2e8b57")` Write a 2-sentence caption for your chart

.pull-left[
.font90[
**1.** What does the chart **show**?

**2.** What is **surprising** or notable?

<br>

Use **precise language** &mdash; exact numbers, clear comparison.
]
]

.pull-right[
.content-box-blue[
.font90[*"Home teams won **58.3%** of games in our dataset, compared to **41.7%** for away teams. The edge is real, but smaller than the roar of a home crowd on TV might suggest."*]
]
]

.font80[`r fontawesome::fa("people-arrows", fill = "#4a7fb5")` &nbsp; Read your caption to your **partner**. Is every number specific and correct?]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-01-introduction-to-research/pm/01-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-01-introduction-to-research/pm/01-pm-datacore.pdf")
to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-01-foundations-of-inquiry/session-01-introduction-to-research/pm/01-pm-datacore.html")
```
