---
title: "DataCORE"
subtitle: "Concept Review, Technical Rehearsal & Summit Ready Check"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | July 15, 2026</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "60%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
import seaborn as sns
sns.set_theme(style="whitegrid")
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.big-stat { font-size: 2.2rem; font-weight: 700; color: #2e8b57; }
.card-grid { display: flex; flex-wrap: wrap; gap: 0.7em; margin-top: 0.6em; }
.card { flex: 1 1 30%; background: #e7f3ec; border-radius: 8px; padding: 0.5em 0.8em; }
.card h4 { margin: 0 0 0.2em 0; color: #2e8b57; font-size: 0.95em; }
.card p { margin: 0; font-size: 0.8em; }
.chev-row-sm { display: flex; align-items: center; gap: 0.8em; margin-bottom: 0.3em; }
.chev-sm { flex: 0 0 50px; height: 56px; background: #cfe8d8;
  clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
  display: flex; align-items: center; justify-content: center;
  font-weight: bold; color: #1a4731; }
.chev-text { flex: 1; }
```


# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Research Concept Review Game <span style="color:#2e8b57;">(1:00&ndash;1:45)</span></h4></li>
    <li><h4>Full Technical Rehearsal + Q&A Prep <span style="color:#2e8b57;">(1:45&ndash;2:45)</span></h4></li>
    <li><h4>Scientific Storytelling &mdash; Summit Ready Check <span style="color:#2e8b57;">(2:45&ndash;3:00)</span></h4></li>
</ol>

.font90[*One last sweep for misconceptions, one final run under Summit conditions, and a check that every box is ticked before the big day.*]

```{python load-data, include=FALSE}
wages = pd.read_csv("wages.csv")
```

---
class: segue-green

# 1:00 &ndash; 1:45 &middot; Research
## .white[Concept Review Game]

---
### `r fontawesome::fa("gamepad", height = "1em", fill = "#2e8b57")` A competitive final scan for misconceptions

.pull-left[
.font90[
**Kahoot** or **Quizlet Live** &mdash; teams compete on key concepts from all 14 sessions:

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("dna", fill = "#2e8b57")`</div><div class="chev-text">Hypothesis vs. theory · types of variables</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("code-branch", fill = "#2e8b57")`</div><div class="chev-text">Correlation vs. causation · p-value meaning</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("quote-right", fill = "#2e8b57")`</div><div class="chev-text">Citation format · research design terms</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("robot", fill = "#2e8b57")`</div><div class="chev-text">AI literacy concepts</div></div>
]
]

.pull-right[
.content-box-yellow[
.font82[`r fontawesome::fa("bullhorn", fill = "#c08a00")` &nbsp; **Instructor watches the scoreboard.** Any concept where class accuracy is low gets a **2-minute clarification on the spot** &mdash; these are exactly the gaps adult guests probe in Summit Q&A.]
]
]

---
### `r fontawesome::fa("circle-check", height = "1em", fill = "#2e8b57")` Concept quick-check &mdash; get these right

<div class="card-grid">
  <div class="card">
    <h4>Correlation &ne; Causation</h4>
    <p>Two things moving together doesn't mean one causes the other.</p>
  </div>
  <div class="card">
    <h4>p-value</h4>
    <p>Probability of data this extreme *if the null were true* &mdash; not "chance the result is real."</p>
  </div>
  <div class="card">
    <h4>Hypothesis vs. Theory</h4>
    <p>A hypothesis is a testable guess; a theory is a well-supported explanation.</p>
  </div>
  <div class="card">
    <h4>Variable types</h4>
    <p>Categorical vs. numeric; independent (cause) vs. dependent (effect).</p>
  </div>
  <div class="card">
    <h4>Citation</h4>
    <p>Every in-text cite appears in References, in one consistent style.</p>
  </div>
  <div class="card">
    <h4>AI literacy</h4>
    <p>AI assists and can be wrong; you verify, and you disclose its use.</p>
  </div>
</div>

---
### `r fontawesome::fa("chart-line", height = "1em", fill = "#2e8b57")` Self-check: what does a correlation *not* prove?

.pull-left[
```{python corr-demo, echo=FALSE, out.width="98%", fig.width=5.2, fig.height=3.4}
import numpy as np
r = np.corrcoef(wages.experience_years, wages.hourly_wage)[0, 1]
fig, ax = plt.subplots()
ax.scatter(wages.experience_years, wages.hourly_wage, alpha=0.5, s=14, color="#2e8b57")
ax.set_title(f"Experience & wage move together (r = {r:.2f})", fontsize=13, fontweight="bold")
ax.set_xlabel("Experience (years)", fontsize=12)
ax.set_ylabel("Hourly wage ($)", fontsize=12)
plt.tight_layout(); plt.show()
```
]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("circle-question", fill = "#c08a00")` &nbsp; A positive **r** tells you experience and wage rise together. It does **not** prove experience *causes* higher wages &mdash; education, age, or the type of job could drive both.]
]
.content-box-green[
.font82[`r fontawesome::fa("check", fill = "#009E73")` &nbsp; In Q&A, say what the number **is** (an association) and what it **isn't** (proof of cause).]
]
]

.center[`r timer(45L)`]

---
class: segue-green

# 1:45 &ndash; 2:45 &middot; Full Technical
## .white[Rehearsal + Q&A Prep]

---
### `r fontawesome::fa("display", height = "1em", fill = "#2e8b57")` One final run &mdash; in Summit conditions

.pull-left[
.font90[
No more practice mode. This run is the real thing:

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("tv", fill = "#2e8b57")`</div><div class="chev-text">Slides up on the <strong>big screen</strong>.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("person", fill = "#2e8b57")`</div><div class="chev-text">Team <strong>stands</strong> and presents, in speaking order.</div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("microphone-lines", fill = "#2e8b57")`</div><div class="chev-text">Instructor plays an <strong>audience member</strong>.</div></div>
]
]

.pull-right[
.content-box-blue[
.font85[`r fontawesome::fa("triangle-exclamation", fill = "#4a7fb5")` &nbsp; Rehearse the **transitions between speakers** &mdash; the awkward hand-off is where most teams lose time and momentum.]
]
]

---
### `r fontawesome::fa("fire", height = "1em", fill = "#2e8b57")` One hard question per team

.font85[The instructor asks the kind of question a sharp guest will ask:]

<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("arrows-rotate", fill = "#2e8b57")`</div><div class="chev-text"><em>"How do you know it isn't the reverse &mdash; that Y causes X?"</em></div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("scale-unbalanced", fill = "#2e8b57")`</div><div class="chev-text"><em>"What would change your conclusion?"</em></div></div>
<div class="chev-row-sm"><div class="chev-sm">`r fontawesome::fa("users-between-lines", fill = "#2e8b57")`</div><div class="chev-text"><em>"Did you test whether this holds for different subgroups?"</em></div></div>

.content-box-yellow[
.font85[`r fontawesome::fa("handshake-angle", fill = "#c08a00")` &nbsp; **When you can't fully answer:** *"That's a great question. Our study didn't address that directly, but based on what we found&hellip;"* A professional, honest response beats a made-up answer.]
]

.center[`r timer(60L)`]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[Summit Ready Check]

---
### `r fontawesome::fa("clipboard-check", height = "1em", fill = "#2e8b57")` The Summit Ready checklist

.pull-left[.font88[
Run it as a class &mdash; every item gets a &check;:

.content-box-green[
- Title slide with **full team names**
- **References** slide present
- **AI attribution** statement included
- **Every chart has a caption**
- All members **know their speaking part**
- Presentation is **under 12 minutes**
]
]]

.pull-right[.font90[
.content-box-yellow[
**`r fontawesome::fa("triangle-exclamation", fill = "#c08a00")` Any unchecked box = emergency fix**

Anything not ticked gets **5 minutes of focused attention** before 3:00 PM. Better to catch it now than on the Summit stage.
]
]]

.center[`r timer(15L)`]

---
class: segue-green

# .white[Rehearsed. Pressure-tested. Ready.]
## .white[Next time you present, it's the Summit.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-13-full-rehearsal-peer-feedback/pm/13-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-03-synthesis-communication-presentation/session-13-full-rehearsal-peer-feedback/pm/13-pm-datacore.pdf")
```
