---
title: "Research Methods"
subtitle: "Day 10: Building Your Presentation — Methods, Results & Citations"
# author: "Hussain Hadah (he/him)"
# date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | `r rmarkdown::metadata$title` | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- `r rmarkdown::metadata$subtitle` | `r format(Sys.time(), '%d %B %Y')`-->

---

class: title-slide
background-position: 10% 90%, 100% 50%
background-size: 160px, 50% 100%
background-color: #0148A4

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = 'authoryear',
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.
options(modelsummary_factory_latex = "kableExtra")
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```


```{r, cache=FALSE, message=FALSE, warning=FALSE, include=TRUE, eval=TRUE, results=FALSE, echo=FALSE, tidy.opts = list(width.cutoff = 50), tidy = TRUE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, grid)

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold',
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

theme_set(hrbrthemes::theme_ipsum() + theme_customs)

# COLOR PALLETES
library(paletteer)

### COLOR BLIND PALLETES
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: auto !important; /* Adjusts table width */
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    .chev-row {
      display: flex;
      align-items: center;
      gap: 0.9em;
      margin-bottom: 0.6em;
    }
    .chev {
      flex: 0 0 62px;
      height: 78px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
      font-weight: bold;
      font-size: 1.7em;
      color: #1a1a1a;
    }
    .chev-text {
      flex: 1;
    }
    .chev-sm {
      flex: 0 0 52px;
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
      display: flex;
      align-items: center;
      justify-content: center;
    }
    .chev-row-sm {
      display: flex;
      align-items: center;
      gap: 0.8em;
      margin-bottom: 0.25em;
    }

    .chev-flow {
      display: flex;
      gap: 0.7em;
      margin-top: 1em;
    }
    .chev-step {
      flex: 1;
      text-align: center;
    }
    .chev-step p {
      text-align: left;
    }
    .chev-right {
      height: 58px;
      background: #c9e4f6;
      clip-path: polygon(0 0, 86% 0, 100% 50%, 86% 100%, 0 100%, 14% 50%);
      display: flex;
      align-items: center;
      justify-content: center;
      margin-bottom: 0.6em;
    }

    .three-col {
      display: flex;
      gap: 1.5em;
      margin-top: 1.5em;
    }
    .three-col .col {
      flex: 1;
      text-align: center;
    }
    .three-col .col p:first-child {
      margin-bottom: 0.4em;
    }

```

# .text-shadow[.white[Outline for Today]]

<ol>
    <li><h4 class="white">The Presentation Outline (Refresher)</h4></li>
    <li><h4 class="white">Goal 1: Finish Your Methods Slide</h4></li>
    <li><h4 class="white">Goal 2: Compile Your Results</h4></li>
    <li><h4 class="white">Goal 3: Enrich Discussion & Conclusion</h4></li>
    <li><h4 class="white">Goal 4: Cite Your Resources</h4></li>
</ol>

---
class: segue-yellow

# Welcome to Day 8! `r top_icon("chalkboard-teacher")`

---
### Today's Goal

.content-box-blue[
.font120[**Today your goal is to work on your slides for your presentation.**]
]

.font105[Let's remember the presentation outline.]

---
### Presentation Outline

.font70[
<div class="chev-row-sm">
  <div class="chev-sm">1</div>
  <div class="chev-text"><strong>Title Slide:</strong> Introduce yourself and the project by naming the title, presenters, date of presentation, the institutions at which the study is conducted. Add logos if you can.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">2</div>
  <div class="chev-text"><strong>Topic Slide:</strong> Name your global challenge, define it (cite your source), state why this is important (cite your sources).</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">3</div>
  <div class="chev-text"><strong>Literature Review Slide:</strong> What do you know about this issue based on your literature review? Have a couple statements arranged by ideas/themes. Cite your sources.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">4</div>
  <div class="chev-text"><strong>Research Question Slide:</strong> Name your research question(s) and if possible hypothesis.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">5</div>
  <div class="chev-text"><strong>Methods Slide:</strong> State that your work is secondary research. Introduce your data source (including the year data was collected) &amp; sample (including its size and key characteristics). Name key variables. Name the tools you used to analyze data.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">6</div>
  <div class="chev-text"><strong>Results Slides:</strong> First share descriptive statistics for each key variable (mean, median, mode, range). Provide numbers, graphs, tables as suitable. This can be multiple slides.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">7</div>
  <div class="chev-text"><strong>Discussion:</strong> How you interpret what you have found.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">8</div>
  <div class="chev-text"><strong>Conclusion:</strong> Final words and good-bye.</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">9</div>
  <div class="chev-text"><strong>References:</strong> List your references. Be consistent with the style.</div>
</div>
]

---
### REMEMBER? Using the 5 W's (and How)

.font90[
<div class="chev-row-sm">
  <div class="chev-sm">`r fontawesome::fa("user", height = "1.3em", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Who</strong> &mdash; who is involved or affected? (teenagers, athletes, teachers, polar bears)</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">`r fontawesome::fa("circle-question", height = "1.3em", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>What</strong> &mdash; what aspect interests you? (mental health, economic cost, performance)</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">`r fontawesome::fa("location-dot", height = "1.3em", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Where</strong> &mdash; a certain location or context? (your city, online communities, lakes)</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">`r fontawesome::fa("clock", height = "1.3em", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>When</strong> &mdash; a time period? (past 5 years, during summer, historical period)</div>
</div>

<div class="chev-row-sm">
  <div class="chev-sm">`r fontawesome::fa("gears", height = "1.3em", fill = "#0148A4")`</div>
  <div class="chev-text"><strong>Why/How</strong> &mdash; causes or processes? (why something happens, how factors interact)</div>
</div>
]

---
class: segue-green

# 1st GOAL: <br> Finish Your Methods Slide `r top_icon("clipboard-list")`

### .white[Make sure it has the following information:]

---
### The Methods Slide Includes:

.font85[
.content-box-blue[
**`r fontawesome::fa("chess", fill = "#4a7fb5")` Research Design**

- **Approach:** Quantitative (Descriptive)
- **Methodology:** Secondary data analysis of a publicly available data set
]

.content-box-blue[
**`r fontawesome::fa("database", fill = "#4a7fb5")` Data Source**

- Name of the data set &bull; years covered / data collected &bull; institution that collected the data
]

.content-box-blue[
**`r fontawesome::fa("users", fill = "#4a7fb5")` Sample**

Introduce your sample including its size, nature, and key characteristics.
]

.content-box-blue[
**`r fontawesome::fa("sliders", fill = "#4a7fb5")` Analyzed Variables**

Simply name each variable you analyzed.
]

.content-box-blue[
**`r fontawesome::fa("toolbox", fill = "#4a7fb5")` Procedure &amp; Tools**

Identify how you approached the data analysis. Steps you took. Tools you used.
]
]

---
class: segue-green

# 2nd GOAL: <br> Compile Your Results `r top_icon("chart-bar")`

### .white[Discuss how to present your results in your presentation.]

---
class: segue-green

# 3rd GOAL: <br> Enrich Your Discussion / Conclusion Slides `r top_icon("comments")`

### .white[Discuss: how do results align with your expectations? Add that to your discussion slide.]

---
### 4th GOAL: Cite Your Resources

.font85[
.content-box-blue[
**`r fontawesome::fa("quote-left", fill = "#4a7fb5")` Throughout the presentation: name utilized resources**

- In research, failing to plan is identified as planning to fail **(Yin &amp; Ogut, 2025)**.
- **Toska (2023)** suggested that counseling centers in T&uuml;rkiye tended to avoid presence on social media platforms.
]

.content-box-green[
**`r fontawesome::fa("list", fill = "#4a7fb5")` At the end of the presentation: full list**

- Toska, G. (2023). *Counseling Centers in Turkiye.* Paper presented at the National Conference on Counseling Psychology. Istanbul, Turkiye.
- Yin &amp; Ogut (2025). *Nexellence ResearchCORE: Unveiling the Power of Inquiry: An Introduction to Research Methods for All Curious Minds.* Nexellence Press. ISBN: 979-8-9988862-1-8
]
]

---
### Homework &mdash; Teamwork

.pull-left[
```{r img12, echo=FALSE, warning=FALSE, fig.width=4, fig.height=6, out.width="55%", fig.show='hold', fig.align='center'}
# Read the image file
img <- readPNG("assets/p12-img.png")

# Plot the image
grid.raster(img)
```
]

.pull-right[
<br>

#### `r fontawesome::fa("list-check", height = "1em", fill = "#4a7fb5")` &nbsp; Finalize your slide content

#### `r fontawesome::fa("people-arrows", height = "1em", fill = "#4a7fb5")` &nbsp; Assign who is presenting which slide

#### `r fontawesome::fa("person-chalkboard", height = "1em", fill = "#4a7fb5")` &nbsp; Practice!
]

---
class: segue

# EXTRA SLIDES FOR DAY 8

---
### SLIDE SHOWDOWN

.pull-left[
.content-box-blue[
.font120[**SHOW US:**]

### `r fontawesome::fa("star", height = "1em", fill = "#4a7fb5")` &nbsp; 2 slides of your best work

### `r fontawesome::fa("circle-question", height = "1em", fill = "#4a7fb5")` &nbsp; 2 slides of your questionable slides
]
]

.pull-right[
```{r img14, echo=FALSE, warning=FALSE, fig.width=4, fig.height=6, out.width="55%", fig.show='hold', fig.align='center'}
# Read the image file
img <- readPNG("assets/p14-img.png")

# Plot the image
grid.raster(img)
```
]

```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
library(renderthis)
# https://hhadah.github.io/nexellence-summer-instit/day-8/08-class.html

to_pdf(from = "~/Documents/GiT/nexellence-summer-instit/day-8/08-class.html",
       to = "~/Documents/GiT/nexellence-summer-instit/day-8/08-class.pdf")
to_pdf(from = "~/Documents/GiT/nexellence-summer-instit/day-8/08-class.html")
```
