---
title: "DataCORE"
subtitle: "Final Analysis, AI Results Check & Portfolio Prep"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | July 10, 2026</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
library(reticulate)
use_virtualenv("r-reticulate", required = TRUE)
suppressMessages(library(tidyverse))
options(readr.show_col_types = FALSE)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "60%")
```

```{python py-setup, include=FALSE}
import pandas as pd
import numpy as np
import matplotlib
matplotlib.use("Agg")
import matplotlib.pyplot as plt
import seaborn as sns
from scipy import stats
import statsmodels.formula.api as smf
sns.set_theme(style="whitegrid")
pd.set_option("display.width", 100)
pd.set_option("display.max_columns", 12)
```

```{r xaringan-panelset, echo=FALSE, eval=TRUE}
xaringanExtra::use_panelset()
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.panelset .panel-tabs .panel-tab { font-size: 0.9em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.big-stat { font-size: 2.2rem; font-weight: 700; color: #2e8b57; }
.card-grid { display: flex; flex-wrap: wrap; gap: 0.7em; margin-top: 0.6em; }
.card { flex: 1 1 30%; background: #e7f3ec; border-radius: 8px; padding: 0.5em 0.8em; }
.card h4 { margin: 0 0 0.2em 0; color: #2e8b57; font-size: 0.95em; }
.card p { margin: 0; font-size: 0.8em; }
.chev-row-sm { display: flex; align-items: center; gap: 0.8em; margin-bottom: 0.3em; }
.chev-sm { flex: 0 0 50px; height: 56px; background: #cfe8d8;
  clip-path: polygon(0 0, 50% 14%, 100% 0, 100% 78%, 50% 100%, 0 78%);
  display: flex; align-items: center; justify-content: center;
  font-weight: bold; color: #1a4731; }
.chev-text { flex: 1; }
```

# .text-shadow[Afternoon: DataCORE]

<ol>
    <li><h4>Final Analysis + AI Results Check</h4></li>
    <li><h4>Final Dataset Documentation &amp; Portfolio Prep</h4></li>
    <li><h4>Scientific Storytelling &mdash; Final Chart Set Selection</h4></li>
</ol>

.font90[*Today you don't learn a new technique &mdash; you **finish.** We'll walk through all three together first; then you have the rest of class to run your analysis, build your portfolio, and lock your charts.*]

```{python load-data, include=FALSE}
wages = pd.read_csv("wages.csv")
```

```{r load-data-r, include=FALSE}
wages <- read.csv("wages.csv")
```

---
class: segue-green

# Final Analysis
## .white[&amp; AI Results Check]

---
### `r fontawesome::fa("arrows-rotate", height = "1em", fill = "#2e8b57")` Run it clean &mdash; exactly once

.pull-left[
.font90[
Before you trust a result, prove it **reproduces**:

<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>Restart</strong> the kernel / session (clear memory).</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text"><strong>Run top to bottom</strong> with no skipping.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text">Confirm <strong>every number</strong> still appears &mdash; no errors, no leftovers.</div></div>
]
]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("triangle-exclamation", fill = "#c08a00")` &nbsp; If it only works when you run cells **out of order**, it isn't a finished analysis &mdash; it's a lucky accident.]
]
]

.font80[**Every code block carries a one-line comment** explaining its purpose &mdash; that's what makes it a portfolio piece.]

---
### `r fontawesome::fa("list-check", height = "1em", fill = "#2e8b57")` A clean, commented pipeline

.panelset[
.panel[.panel-name[Python]
```{python}
wages = pd.read_csv("wages.csv")             # load the raw data
wages = wages.dropna()                       # drop incomplete rows
gap = (wages.loc[wages.gender=="Male","hourly_wage"].mean()
       - wages.loc[wages.gender=="Female","hourly_wage"].mean())   # gender wage gap
r, p = stats.pearsonr(wages.experience_years, wages.hourly_wage)   # experience-wage link
print(f"N = {len(wages)}, gap = ${gap:.2f}, r = {r:.2f} (p = {p:.4f})")
```
]
.panel[.panel-name[R]
```{r}
wages <- read.csv("wages.csv")               # load the raw data
wages <- na.omit(wages)                       # drop incomplete rows
gap <- mean(wages$hourly_wage[wages$gender=="Male"]) -
       mean(wages$hourly_wage[wages$gender=="Female"])   # gender wage gap
ct <- cor.test(wages$experience_years, wages$hourly_wage) # experience-wage link
cat(sprintf("N = %d, gap = $%.2f, r = %.2f (p = %.4f)\n",
            nrow(wages), gap, ct$estimate, ct$p.value))
```
]
]

.font80[`r fontawesome::fa("comment", fill = "#4a7fb5")` &nbsp; Notice: **one comment per line of intent.** A stranger (or future you) can read it top to bottom and follow the logic.]

---
### `r fontawesome::fa("robot", height = "1em", fill = "#2e8b57")` AI results check &mdash; an honest second opinion

.pull-left[
.content-box-blue[
.font82[
**Paste your key results into the AI and ask:**

*"Are there any obvious errors or red flags in these results?"*
]
]
.font80[Then compare its critique to **your own judgment.**]
]

.pull-right[
.content-box-green[
.font80[
**The AI is good at catching:**
- Impossible values (a negative age, r &gt; 1)
- Mismatched units or decimal slips
- Claims stronger than the numbers support
]
]
.content-box-yellow[
.font80[
**It can miss:**
- Context only *you* know (how data was collected)
- Whether the test fits the question
]
]
]

<div style="clear: both;"></div>

.font80[`r fontawesome::fa("scale-balanced", fill = "#4a7fb5")` &nbsp; **Does it flag anything you missed? Does it miss anything important?** You are still the analyst &mdash; the AI is the reviewer.]

---
### `r fontawesome::fa("vials", height = "1em", fill = "#2e8b57")` Robustness: drop a control, re-run

.panelset[
.panel[.panel-name[Python]
```{python}
full = smf.ols("hourly_wage ~ experience_years + C(education) + C(gender)",
               data=wages).fit()                                   # full model
reduced = smf.ols("hourly_wage ~ experience_years + C(gender)",
                  data=wages).fit()                                # drop education
print(f"experience coef  (full)    = {full.params['experience_years']:.3f}")
print(f"experience coef  (reduced) = {reduced.params['experience_years']:.3f}")
```
]
.panel[.panel-name[R]
```{r}
full <- lm(hourly_wage ~ experience_years + education + gender, data = wages) # full
reduced <- lm(hourly_wage ~ experience_years + gender, data = wages)  # drop education
cat(sprintf("experience coef (full)    = %.3f\n", coef(full)["experience_years"]))
cat(sprintf("experience coef (reduced) = %.3f\n", coef(reduced)["experience_years"]))
```
]
]

```{python rb-extract, include=FALSE}
b_full = float(full.params["experience_years"])
b_red  = float(reduced.params["experience_years"])
pct = abs(b_red - b_full) / abs(b_full) * 100
```

.content-box-yellow[
.font82[`r fontawesome::fa("magnifying-glass", fill = "#c08a00")` &nbsp; Dropping **education** moves the experience coefficient by about **`r sprintf("%.0f", py$pct)`%**. If a control swings your main coefficient a lot, that sensitivity **is a limitation** &mdash; document it honestly.]
]

---
class: segue-green

# Portfolio Prep
## .white[Your Analysis as a Research Artifact]

---
### `r fontawesome::fa("box-open", height = "1em", fill = "#2e8b57")` Your portfolio is an artifact, not a folder of files

<div class="card-grid">
  <div class="card">
    <h4>`r fontawesome::fa("file-code", fill = "#2e8b57")` Commented notebook</h4>
    <p>Every code block has a one-line comment explaining its purpose.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("image", fill = "#2e8b57")` Exported figures</h4>
    <p>PNG, 300 DPI, descriptive file names.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("table", fill = "#2e8b57")` Summary table</h4>
    <p>Formatted statistics table with an APA caption.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("broom", fill = "#2e8b57")` Data cleaning log</h4>
    <p>What you changed, why, and what you chose <em>not</em> to change.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("robot", fill = "#2e8b57")` AI usage log</h4>
    <p>Tools used, prompts that worked, errors the AI made.</p>
  </div>
  <div class="card">
    <h4>`r fontawesome::fa("file-pdf", fill = "#2e8b57")` Exported PDF</h4>
    <p>The notebook as a PDF &mdash; what you display at the Summit.</p>
  </div>
</div>

---
### `r fontawesome::fa("table", height = "1em", fill = "#2e8b57")` Summary statistics table + caption

.panelset[
.panel[.panel-name[Python]
```{python}
summary = wages.groupby("education")["hourly_wage"].agg(
    ["count", "mean", "median", "std"]).round(2)   # summary stats by group
print(summary)
```
]
.panel[.panel-name[R]
```{r}
wages |> group_by(education) |>
  summarise(count = n(), mean = round(mean(hourly_wage),2),
            median = median(hourly_wage), std = round(sd(hourly_wage),2))
```
]
]

.content-box-green[
.font85[`r fontawesome::fa("quote-left", fill = "#2e8b57")` &nbsp; **Table 1.** *Hourly wage by education level (N = 220).* **Note.** Data from the 2023 wage survey; values in U.S. dollars per hour.]
]

---
### `r fontawesome::fa("clipboard-list", height = "1em", fill = "#2e8b57")` Two logs that make you credible

.pull-left[
.content-box-blue[
.font80[
**`r fontawesome::fa("broom", fill = "#4a7fb5")` Data cleaning log**

| Step | Why |
|---|---|
| Dropped 6 blank rows | No wage recorded |
| Recoded "M/F" &rarr; gender | Consistent labels |
| Kept 3 high wages | Real, not errors |
]
]]

.pull-right[
.content-box-yellow[
.font80[
**`r fontawesome::fa("robot", fill = "#c08a00")` AI usage log**

| Prompt | Outcome |
|---|---|
| "Explain r = 0.42" | Useful plain-English |
| "Write my groupby" | Worked after 1 fix |
| "Is this causal?" | Over-claimed &mdash; corrected |
]
]]

<div style="clear: both;"></div>

.font82[`r fontawesome::fa("circle-check", fill = "#4a7fb5")` &nbsp; Logging what you changed **and** how you used AI is honesty, not bureaucracy &mdash; it lets anyone retrace your steps.]

---
### `r fontawesome::fa("floppy-disk", height = "1em", fill = "#2e8b57")` Export figures &amp; notebook

.panelset[
.panel[.panel-name[Figure · Python]
```{python, eval=FALSE}
plt.savefig("fig1_experience_wage_scatter.png",
            dpi=300, bbox_inches="tight")     # 300 DPI, descriptive name
```
]
.panel[.panel-name[Figure · R]
```{r, eval=FALSE}
ggsave("fig1_experience_wage_scatter.png",
       width = 6, height = 4, dpi = 300)        # 300 DPI, sized in inches
```
]
.panel[.panel-name[Notebook PDF]
```{r, eval=FALSE}
# Jupyter:   File > Download as > PDF,  or:
#   jupyter nbconvert --to pdf analysis.ipynb
# R Markdown / Quarto:
#   quarto render analysis.qmd --to pdf
```
]
]

.content-box-blue[
.font85[`r fontawesome::fa("file-pdf", fill = "#4a7fb5")` &nbsp; The **exported PDF** is the deliverable you display at the Research Summit &mdash; code, output, figures, and your comments in one clean document.]
]

---
class: segue-green

# Scientific Storytelling
## .white[Final Chart Set Selection]

---
### `r fontawesome::fa("filter", height = "1em", fill = "#2e8b57")` Lock your 2&ndash;3 Summit charts

.font90[For each candidate chart, ask three questions. If it fails one &mdash; **simplify or cut it.**]

<div class="three-col">
  <div class="col">
    <div class="content-box-blue">
    <p><strong>`r fontawesome::fa("bullseye", fill="#4a7fb5")` 1. Relevant?</strong></p>
    <p class="font80">Does it <strong>directly answer</strong> your research question?</p>
    </div>
  </div>
  <div class="col">
    <div class="content-box-green">
    <p><strong>`r fontawesome::fa("eye", fill="#009E73")` 2. Readable?</strong></p>
    <p class="font80">Is it legible at <strong>presentation size</strong>, from the back of the room?</p>
    </div>
  </div>
  <div class="col">
    <div class="content-box-yellow">
    <p><strong>`r fontawesome::fa("stopwatch", fill="#c08a00")` 3. Explainable?</strong></p>
    <p class="font80">Can you explain it in <strong>under 30 seconds</strong>?</p>
    </div>
  </div>
</div>

--

.content-box-blue[
.font90[`r fontawesome::fa("award", fill = "#4a7fb5")` &nbsp; **A clear, simple chart beats a complex one every time.** When in doubt, cut a variable, not a viewer.]
]

---
class: segue-green

# Your Turn
## .white[Work for the Rest of Class]

---
### `r fontawesome::fa("list-check", height = "1em", fill = "#2e8b57")` What to work on now

.content-box-blue[
.font90[We've covered all three tasks. Use the rest of class to **finish** &mdash; work in your team, and wave me over when you need help.]
]

.font78[
<div class="chev-row-sm"><div class="chev-sm">1</div><div class="chev-text"><strong>Final analysis + AI results check</strong> &mdash; restart, run top-to-bottom once, confirm every number reproduces with a comment on each block. Paste your key results into the AI for an error check, then run one robustness check (drop a control, re-run) and note any big swing.</div></div>
<div class="chev-row-sm"><div class="chev-sm">2</div><div class="chev-text"><strong>Portfolio prep</strong> &mdash; assemble the artifact: commented notebook, exported figures (300 DPI), summary table + APA caption, data-cleaning log, AI-usage log, and the exported PDF you'll show at the Summit.</div></div>
<div class="chev-row-sm"><div class="chev-sm">3</div><div class="chev-text"><strong>Final chart set</strong> &mdash; lock your 2&ndash;3 Summit charts. Each must be relevant, readable at presentation size, and explainable in under 30 seconds.</div></div>
]

.center[`r timer(90L)`]

---
class: segue-green

# .white[One clean run. An honest check. A portfolio you'd show anyone.]
## .white[That's what turns analysis into research.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)

to_pdf(from = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-02-investigation-and-analysis/session-10-synthesizing-findings/pm/10-pm-datacore.html",
       to = "/Users/hhadah/Projects/GiT/nexellence-summer-instit/week-02-investigation-and-analysis/session-10-synthesizing-findings/pm/10-pm-datacore.pdf")
```
