---
title: "DataCORE"
subtitle: "Scientific Storytelling: From Literature Review to Results Language"
# author: "Hussain Hadah (he/him)"
output:
  xaringan::moon_reader:
    css: ["default", "assets/tulane-fonts.css", "assets/tulane.css"]
    self_contained: false
    seal: true
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      navigation:
        scroll: false
---
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">Nexellence Institute | DataCORE | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- footer -->

---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE,
               fig.align = "center", dpi = 150, fig.width = 6, fig.height = 3.4,
               out.width = "62%")
```

```{r countdown-setup, echo=FALSE}
library(countdown)
timer <- function(minutes = 1L, seconds = 0L, ...) {
  countdown::countdown(
    minutes = minutes, seconds = seconds,
    style = "position: relative; display: inline-block; margin: 0.4em auto;",
    color_border = "#2e8b57",
    color_running_background = "#2e8b57",
    color_running_text = "#FFFFFF",
    color_finished_background = "#D55E00",
    font_size = "2rem",
    warn_when = 30L,
    ...
  )
}
```

```{css, echo=FALSE, eval=TRUE}
.remark-code, .remark-inline-code { font-size: 0.72em; }
.three-col { display: flex; gap: 1.5em; margin-top: 1em; }
.three-col .col { flex: 1; }
.mock-slide { border: 2px solid #4a7fb5; border-radius: 8px; padding: 0.6em 0.9em;
              background: #ffffff; box-shadow: 2px 2px 6px rgba(0,0,0,0.15); }
.mock-slide h4 { margin: 0 0 0.3em 0; color: #0148A4; }
```

### `r fontawesome::fa("table-cells", height = "1em", fill = "#2e8b57")` Warm-up: creating a correlation matrix

.font85[A **correlation matrix** shows the pairwise correlation between every pair of numeric variables &mdash; usually your **first look** at how the variables in your data move together.]

.pull-left[
.font85[**R** &mdash; `cor()` is built into base R (the `stats` package, loaded automatically &mdash; nothing to install):]

```{r cor-matrix-r, eval=TRUE}
wages <- read.csv("wages.csv")
cor(wages[, c("experience_years",
              "hourly_wage")])
```
]

.pull-right[
.font85[**Python** &mdash; needs the **pandas** library (`pip install pandas`):]

```python
import pandas as pd

wages = pd.read_csv("wages.csv")
wages[["experience_years",
       "hourly_wage"]].corr()
```
]

.content-box-blue[
.font80[`r fontawesome::fa("palette", fill = "#4a7fb5")` &nbsp; **Hard to read as a table of numbers?** Turn it into a **heatmap** &mdash; next slide.]
]

---
### `r fontawesome::fa("palette", height = "1em", fill = "#2e8b57")` Warm-up: the correlation matrix as a heatmap

.font85[A **heatmap** colors each cell by the strength of the correlation &mdash; patterns jump out that a table of numbers hides.]

.pull-left[
.font85[**R** &mdash; needs the **corrplot** library (`install.packages("corrplot")`):]

```{r cor-heatmap-r, eval=TRUE, fig.width=4.5, fig.height=4.5, out.width="38%"}
library(corrplot)

vars <- wages[, c("experience_years",
                  "hourly_wage")]
colnames(vars) <- c("Years of experience",
                    "Hourly wage")

corrplot(cor(vars), method = "color",
         addCoef.col = "black")
```
]

.pull-right[
.font85[**Python** &mdash; needs **seaborn** + **matplotlib** (`pip install seaborn matplotlib`):]

```python
import seaborn as sns
import matplotlib.pyplot as plt

corr = wages.rename(columns={
    "experience_years": "Years of experience",
    "hourly_wage": "Hourly wage"
}).corr(numeric_only=True)

sns.heatmap(corr, annot=True,
            cmap="coolwarm", vmin=-1, vmax=1)
plt.show()
```

.content-box-blue[
.font80[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; `annot=True` / `addCoef.col` print the numbers **on** the colors &mdash; best of both worlds.]
]
]

---

# .text-shadow[Afternoon: Scientific Storytelling]

<ol>
    <li><h4>The Setup &mdash; Your Literature Review as a Story <span style="color:#2e8b57;">(1:15&ndash;1:45)</span></h4></li>
    <li><h4>Practice &mdash; Your Two Slides, Out Loud <span style="color:#2e8b57;">(1:45&ndash;2:45)</span></h4></li>
    <li><h4>The Payoff &mdash; Results Language <span style="color:#2e8b57;">(2:45&ndash;3:00)</span></h4></li>
</ol>

.font90[*This morning you learned the **rules** of academic writing. This afternoon you learn the **craft**: turning your literature review into two slides an audience can follow &mdash; and your findings into one sentence they can act on.*]

---
class: segue-green

# 1:00 &ndash; 1:45 &middot; The Setup
## .white[Your Literature Review as a Story]

---
### `r fontawesome::fa("book-open", height = "1em", fill = "#2e8b57")` Why storytelling? Because nobody remembers a list

.pull-left[
.font90[
Every good research talk is a **story** with the same beats:

1. **Setting** &mdash; here's the world and what we know about it *(background)*
2. **Tension** &mdash; but here's what we **don't** know *(the gap)*
3. **Quest** &mdash; so we asked this question *(research question)*
4. **Resolution** &mdash; and here's what we found *(results)*
]
]

.pull-right[
.content-box-blue[
.font85[`r fontawesome::fa("lightbulb", fill = "#4a7fb5")` &nbsp; Your audience won't remember your fourth citation. They **will** remember the tension: *"we know A and B &mdash; but nobody knows C, so we asked&hellip;"*]
]
]

.font85[Today we practice the two moments where storytelling matters most: the **setup** (literature &rarr; gap &rarr; question) and the **payoff** (results).]

---
### `r fontawesome::fa("filter", height = "1em", fill = "#2e8b57")` The funnel: a chapter becomes two slides

.pull-left[
.font90[
Your literature review is **pages** long. Your presentation gets **two slides**:

.content-box-blue[
.font85[**Slide 1 &mdash; Background**<br>
Key **definitions** + **2&ndash;3 cited findings**. What the audience must know to care about your question.]
]

.content-box-green[
.font85[**Slide 2 &mdash; Gap + Research Question**<br>
What's **missing** from what we know &mdash; and the **question** that follows from it.]
]
]
]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("compress", fill = "#c08a00")` &nbsp; This is not "shrinking" your literature review &mdash; it's **distilling** it. You keep only what sets up the question. Everything else stays in the written report.]
]
]

---
### `r fontawesome::fa("layer-group", height = "1em", fill = "#2e8b57")` Anatomy of the Background slide

.pull-left[
.font88[
**1. Key definitions (1&ndash;2 terms)**
- Only terms your audience **needs** &mdash; not a glossary.
- One line each, in your own words.

**2. What we know (2&ndash;3 cited findings)**
- Organized by **theme**, not by author.
- One finding per line &mdash; **no paragraphs**.
- Each ends with a citation: **(Author, Year)**.
]
]

.pull-right[
.content-box-red[
.font85[`r fontawesome::fa("triangle-exclamation", fill = "#fff")` &nbsp; **The #1 mistake:** pasting sentences from your written review onto the slide. A slide is a **billboard**, not a page. If it takes more than 5 seconds to read a line, cut it.]
]
]

---
### `r fontawesome::fa("eye", height = "1em", fill = "#2e8b57")` Worked example &mdash; the wildfire PBL case

.font80[From this morning's climate case: *Does temperature rise affect wildfire frequency?*]

.mock-slide[
#### Background: Temperature & Wildfires
.font80[
**Key terms:** &nbsp;*Acres burned* = total area destroyed by wildfire per year &nbsp;&middot;&nbsp; *Fire season* = months when large fires occur

**What we know:**
- Global average temperature has risen about 1.1&deg;C since pre-industrial times (IPCC, 2021)
- Fire seasons in the western U.S. have grown longer and more severe since the 1980s (Westerling, 2016)
- Drought and forest management also shape fire risk &mdash; temperature is not the only factor (Abatzoglou & Williams, 2016)
]
]

.font80[`r fontawesome::fa("check", fill = "#2e8b57")` &nbsp; Two terms defined. Three findings, one line each, each cited. A stranger can read this slide in **20 seconds** and be ready for your question.]

---
### `r fontawesome::fa("puzzle-piece", height = "1em", fill = "#2e8b57")` Anatomy of the Gap + Research Question slide

.pull-left[
.font88[
**1. The gap (1&ndash;2 sentences)**
- What do the studies you cited **not** answer?
- Point back at your Background slide: the gap should feel **inevitable** after it.

**2. The research question (verbatim)**
- Word-for-word the question your team wrote &mdash; don't paraphrase it live.
- Add your **hypothesis** if you have one.
]
]

.pull-right[
.content-box-blue[
.font85[`r fontawesome::fa("link", fill = "#4a7fb5")` &nbsp; The magic sentence connecting the two slides:

*"Because we don't yet know **[gap]**, we ask: **[research question]**."*

If that sentence doesn't flow, one of the two slides needs work.]
]
]

---
### `r fontawesome::fa("eye", height = "1em", fill = "#2e8b57")` Worked example &mdash; Gap + Research Question

.mock-slide[
#### The Gap & Our Question
.font80[
**What's missing:** Most studies track how **long** fire seasons have become. Fewer connect temperature **directly** to how much land actually burns each year over recent decades.

**Our research question:** *Is average summer temperature associated with annual acres burned in the United States, 1990&ndash;2020?*

**Our hypothesis:** Years with hotter summers will show more acres burned.
]
]

--

.font85[Say the magic sentence out loud: *"Because we don't yet know whether temperature tracks **acres burned**, we ask: is average summer temperature associated with annual acres burned?"* `r fontawesome::fa("arrow-right", fill="#2e8b57")` It flows &mdash; the slides are doing their job.]

---
### `r fontawesome::fa("ban", height = "1em", fill = "#2e8b57")` Four ways these slides go wrong

.three-col[
.col[
.content-box-red[
.font78[
**The wall of text**<br>
Full sentences from the written review. Audience reads instead of listens.
]
]
.content-box-red[
.font78[
**The name-drop**<br>
"Smith (2020) found&hellip; Jones (2021) found&hellip;" &mdash; organized by author, no theme, no point.
]
]
]
.col[
.content-box-red[
.font78[
**The orphan question**<br>
A research question that doesn't follow from the gap &mdash; the audience wonders *"why THIS question?"*
]
]
.content-box-red[
.font78[
**The missing citation**<br>
A confident claim with no (Author, Year). Instantly costs you credibility.
]
]
]
.col[
.content-box-green[
.font80[
`r fontawesome::fa("circle-check", fill = "#2e8b57")` **The fix is the same for all four:**

One line per idea. Themes, not authors. Gap points at the question. Every claim cited.
]
]
]
]

---
class: segue-green

# `r fontawesome::fa("trophy", fill = "#FFFFFF")` 1:45 &ndash; 2:45 &middot; Team Task
## .white[Your Literature Review &rarr; Two Slides]

---
### `r fontawesome::fa("flag-checkered", height = "1em", fill = "#2e8b57")` Step 1 &mdash; Convert your literature review &mdash; ~25 min

.pull-left[
.font90[
With your team, convert your literature review into the **two slides**:

1. **Background slide** &mdash; 1&ndash;2 key definitions + **2&ndash;3 cited findings**, one line each, by theme.
2. **Gap + Research Question slide** &mdash; the gap in 1&ndash;2 sentences, then your RQ **verbatim** (+ hypothesis).
3. Test with the **magic sentence**: *"Because we don't yet know \_\_\_, we ask \_\_\_."*
]
]

.pull-right[
.content-box-blue[
.font85[
`r fontawesome::fa("circle-check", fill = "#4a7fb5")` **Quality bar:**

- Could a stranger read each slide in **20 seconds**?
- Is every claim **cited** (Author, Year)?
- Does the question feel **inevitable** after the gap?
]
]

.center[`r timer(25L)`]
]

---
### `r fontawesome::fa("people-arrows", height = "1em", fill = "#2e8b57")` Step 2 &mdash; Present to a partner &mdash; 10 min

.pull-left[
.font90[
Pair up with someone from **another team**. Each of you:

1. Present your **two slides** in &le; 2 minutes &mdash; *speak to your partner, don't read the slides.*
2. Partner gives **two pieces of feedback**:
  - `r fontawesome::fa("bolt", fill = "#2e8b57")` *"The clearest moment was&hellip;"*
  - `r fontawesome::fa("lightbulb", fill = "#c08a00")` *"I got lost when&hellip;"*
3. Swap roles.
]
]

.pull-right[
.content-box-yellow[
.font85[`r fontawesome::fa("ear-listen", fill = "#c08a00")` &nbsp; **Listeners:** you are the audience test. If you can't repeat back the gap and the question after 2 minutes, say so &mdash; that's the most useful feedback your partner will get today.]
]

.center[`r timer(10L)`]
]

--

.font85[`r fontawesome::fa("rotate", fill = "#4a7fb5")` &nbsp; Back with your team: apply the feedback to the slides **now**, while it's fresh. These two slides go straight into your final presentation.]

---
class: segue-green

# 2:45 &ndash; 3:00 &middot; Scientific Storytelling
## .white[The Payoff: Results Language]

---
### `r fontawesome::fa("feather-pointed", height = "1em", fill = "#2e8b57")` Two registers &mdash; you need **both**

.pull-left[
.content-box-blue[
.font85[
**The number** (for credibility)<br>
*"The regression coefficient on experience was 0.34 (p &lt; 0.05)."*

Other researchers can **check** this.
]
]
]

.pull-right[
.content-box-green[
.font85[
**The plain language** (for the audience)<br>
*"Our analysis found that workers earned about 34 cents more per hour for each additional year of experience."*

A policymaker can **act** on this.
]
]
]

.content-box-yellow[
.font85[`r fontawesome::fa("scale-balanced", fill = "#c08a00")` &nbsp; Strong results writing puts the **plain sentence first**, then backs it with the numbers in parentheses. Never make the reader decode statistics to learn what you found.]
]

.font80[`r fontawesome::fa("circle-info", fill = "#4a7fb5")` &nbsp; You'll learn to **produce** numbers like these in Sessions 8&ndash;9. Today you learn how to **say** them.]

---
### `r fontawesome::fa("language", height = "1em", fill = "#2e8b57")` Translation practice &mdash; say it both ways

.font85[Each statistical sentence below is **true but unusable** on its own. Translate it out loud:]

.font85[**1.** *"Mean difference = \$2.28, p = 0.02."*]

--

.content-box-green[.font80[`r fontawesome::fa("arrow-right", fill="#2e8b57")` *"Our analysis found that men in our sample earned about **\$2.28 more per hour** than women &mdash; a gap larger than we'd expect from chance alone (p = 0.02)."*]]

.font85[**2.** *"r = 0.62 between summer temperature and acres burned."*]

--

.content-box-green[.font80[`r fontawesome::fa("arrow-right", fill="#2e8b57")` *"Our analysis found that **hotter summers went hand in hand with more land burned** &mdash; a fairly strong association (r = 0.62). Association, not proof of cause."*]]

--

.content-box-blue[.font85[`r fontawesome::fa("magnifying-glass", fill = "#4a7fb5")` &nbsp; Notice the pattern: **plain claim first** (direction + size in real units), **statistics in parentheses**, **caveat if needed**.]]

---
### `r fontawesome::fa("pen-fancy", height = "1em", fill = "#2e8b57")` Draft your team's headline finding &mdash; now

.pull-left[
.font90[
Write **one results sentence** for your team's main finding, in this shape:

.content-box-blue[
.font80[*"Our analysis found that **[plain-English finding]** (**[stat]** = \_\_)."*]
]

Then read it to a partner who is **not** on your team. Did they get it?

.font80[*No statistics yet? Use your strongest **descriptive** finding &mdash; a difference, a trend, a pattern &mdash; and put its number in the parentheses.*]
]
]

.pull-right[
.content-box-yellow[
.font85[
`r fontawesome::fa("circle-check", fill = "#c08a00")` **Checklist:**

- Plain-English claim **first**?
- Direction clear (more / less / higher)?
- The number **with units**?
- Statistics in parentheses?
- No causal verb unless you can defend it?
]
]

.center[`r timer(10L)`]
]

---
class: segue-green

# .white[A research story earns trust twice:]
## .white[once with the setup &mdash; and once with the numbers.]


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
library(renderthis)
to_pdf(from = "07-pm-datacore.html", to = "07-pm-datacore.pdf")
```
